---
title: "Risk-based Weight of Evidence Assessment of the Elasmobranch Resources of Western Australia"
author: "Matias Braccini & Alex Hesp"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  officedown::rdocx_document  
bibliography: 'C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Analyses/Build-Documents/My Library.bib'
csl: 'C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Analyses/Build-Documents/canadian-journal-of-fisheries-and-aquatic-sciences.csl'
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Reports/WoE Advice/2023") })


---



# Background
Over 100 elasmobranch species have been identified in the catches of commercial and/or recreational fisheries in WA. Gummy, dusky, whiskery, and sandbar sharks are the most commonly captured species (~80% of the elasmobranch catch). These species have been targeted in the Temperate Demersal Gillnet and Demersal Longline Fisheries (TDGDLF), which operate in the West Coast and South Coast Bioregions and comprise the West Coast Demersal Gillnet and Demersal Longline (Interim) Managed Fishery (WCDGDLF), which operates between 26° and 33° S, and the Joint Authority Southern Demersal Gillnet and Demersal Longline Managed Fishery (JASDGDLF), which operates from 33° S to the WA/South Australia (SA) border. Sandbar shark and to a lesser extent dusky shark were also targeted by the Northern Shark Fisheries, which operated between 114° 06' E and the WA/Northern Territory (NT) border and ceased operations in 2008-09. The current catch of sharks and rays in other WA commercial fisheries is negligible (< 10 t per annum) and recreational fishers retain very small numbers of sharks in WA. Indigenous catches of these species are considered to be negligible both in terms of historical and extant harvest.

Based on their importance to fisheries, inherent vulnerabilities and sustainability risk, whiskery, gummy, dusky and sandbar sharks have been selected as indicators for the status of the temperate elasmobranch (sharks and rays) ‘suite’ as they represent the range of life history strategies of other elasmobranch species caught by these fisheries. However, with the increasing number of elasmobranch species being listed in national (Environment Protection and Biodiversity Conservation Act 1999, EBPC Act) and international (Convention on International Trade in Endangered Species, CITES; Convention on the Conservation of Migratory Species, CMS) protection lists, species-specific scientific advice on stock status at the local level is needed to comply with national and international legislation.  

For whiskery shark, the initial historical stock assessment was based on a bespoke age- and sex-structured model (i.e., L5 assessment, [@Simpfendorfer2000a]). This assessment determined that female mature biomass in 1997-98 (the last year of the assessed data series) was at 23% of virgin levels, and to achieve management biomass targets with a 50% probability by 2010, recommended catches were required to be maintained below 190 tonnes. For sandbar and dusky sharks, previous stock assessments were based on demographic and mark recapture analyses [@McAuley2007b]. These assessments determined that the combined fishing pressure from WA shark fisheries and cryptic mortality from other fisheries was deemed unsustainable at the effort peak (1990s-2000s). More recently, the weight of evidence assessments for the four indicator species (L1 and L5 assessments for whiskery shark, L1 assessments for gummy, dusky and sandbar sharks), determined that the risk posed by fishing to the sustainability of these stocks was assessed as Medium for 2015-16 (last year of the data series) and for the projected period to 2020-21 [@bracciniRiskbasedWeightEvidence2021]. Prior to the current study, there has been no stock assessments for other shark and ray species captured in WA.  

Following a comprehensive reconstruction of the total catch trajectories of all species for which data are available [@Braccini2021; @bracciniQuantifyingSharkRay2021; @bracciniRecreationalHarvestSharks2021], a range of assessment methods was used to provide species-specific scientific advice on the stock status of all elasmobranch species. First, a Productivity Susceptibility Analysis (PSA) was used to filter out species with historically very low catch volumes and those with low vulnerability  from further analysis. Second, species ranked by the PSA as having high vulnerability were assessed using a combination of spatio-temporal changes in reported catch and effort, catch-only models, State-space surplus production models, and age- and sex-structured integrated models, depending on data availability (Figure 1). Biomass estimates from the production and integrated models were compared to reference levels relating to estimates of Maximum Sustainable Yield (MSY). The estimated biomass expected to achieve MSY (B~MSY~) was considered as the Threshold reference level (B~Thr~) for the stock; 0.5 B~MSY~ was set as the Limit reference level (B~Lim~) and 1.2 B~MSY~ was set as the Target reference level (B~Tar~). A flowchart explaining the steps undertaken in this assessment is shown in Figure 2.

For the indicator species, the major performance measures for assessing stock status are spawning stock levels and fishing mortality (F), estimated using an integrated age-structured assessment model. For the non-indicator species, a range of performance measures for assessing stock status are used based on data availability.

For gummy shark, the estimated biomass and F trajectories showed considerable uncertainty. Biomass trajectories considerably declined in the 1980s following increased exploitation but median biomass remained above target levels throughout the time series. Projections to 2026-27 show a slight increasing trend. For 2021-22 and 2026-27, the probability that the relative biomass was above the limit reference level was 100% and the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Low**. On the basis of this evidence, gummy shark is classified as a **Sustainable stock**.

For dusky shark, biomass trajectories considerably declined in the late 1970s following increased exploitation, with median biomass falling below the threshold reference level in the 2010s, and increasing in recent years following a reduction in catches, with median biomass reaching threshold levels in 2020-21. Projections to 2026-27 show a slight increasing trend. For 2021-22 and 2026-27, the probability that relative biomass was above the limit reference level was 100%. However, the probability that relative biomass was between the limit and threshold reference levels was 47% for 2021-22, and 43% for 2026-27. F trajectories remained at a negligible level at the start of the time series, peaking in the late 1980s and 1990s (when limit reference levels were exceeded), and declining in recent years. For 2021-22, the median F was just below the target reference level. Projections to 2026-27 showed a stable trend. For 2021-22 and 2026-27, the probability that F was below the limit reference level was 100%. However, the probability that F was between the threshold and target levels was 31% for 2026-27. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **High** mostly due to estimation uncertainty as the integrated model was fitted to length composition data only. On the basis of this evidence, dusky shark is classified as a **Recovering stock**.

For whiskery shark, biomass trajectories considerably declined in the late 1940s following increased exploitation, with median biomass falling below the limit reference level between the 1990s and early 2010s, before increasing in recent years following a reduction in catches, with median biomass remaining above target reference levels since 2020-21. Projections to 2026-27 show an increasing trend. For 2021-22 and 2026-27, the probability that relative biomass was above the limit reference level was 87% and 97%, respectively. However, the probability that relative biomass was below the limit, and between the limit and threshold reference levels was 13% and 19%, respectively, for 2021-22. F trajectories remained at a negligible level at the start of the time series, peaking in the 1990s (when limit reference levels were exceeded), and declining in recent years. For 2021-22, median F was considerably below the target reference level. Projections to 2026-27 show a declining trend. For 2021-22 and 2026-27, the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Medium**. On the basis of this evidence, whiskery shark is classified as a **Sustainable stock**.

For sandbar shark, biomass trajectories considerably declined in the 1970s following initial exploitation, with median biomass falling between target and threshold reference levels in the mid 2000s but increasing in recent years following a reduction in catches, with median biomass remaining above target reference levels since the early 2010s. Projections to 2026-27 show an increasing trend. F trajectories remained at a negligible level at the start of the time series, peaking in the late 2000s (when limit reference levels were exceeded) and declining in recent years. For 2021-22, median F was considerably below the target reference level.  Projections to 2026-27 show a stable trend in F. For 2021-22 and 2026-27, the probability that the relative biomass was above the limit reference level was 100% and the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Low**. On the basis of this evidence, sandbar shark is classified as a **Sustainable stock**.

For the non-indicator species, based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **High** for angel sharks, **Medium** for sawsharks, copper, and spinner sharks, and **Low** for the other non-indicator species. For copper and angel sharks, most of the catch has been taken by fisheries outside the jurisdiction of WA and angel sharks were classed as **High** risk due to the high levels of uncertainty in the MSY value inferred from catch-only modelling. On the basis of this evidence, all non-indicator species are classified as **Sustainable stocks**.


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Available time series by species.tiff',sep='')

```
![Figure 1. Available time series for the species assessed as having High vulnerability to fishing by the PSA. F, empirical fishing mortality estimates for the TDGDLF.](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.an,'/Stock assessment flow chart.tiff',sep='')

```
![Figure 2. Flowchart describing the steps undertaken in the current assessment. The number of species assessed by the different assessment methods is shown in brackets.](`r path`){width=800px}






# Executive summary
Over a hundred elasmobranch species have been identified in the catch of commercial and/or recreational fisheries in WA. Gummy, dusky, whiskery and sandbar sharks are the most commonly captured species (~80% of the elasmobranch catch). These species have been targeted in the Temperate Demersal Gillnet and Demersal Longline Fisheries (TDGDLF), which operate in the West Coast and South Coast Bioregions and comprise the West Coast Demersal Gillnet and Demersal Longline (Interim) Managed Fishery (WCDGDLF), which operates between 26° and 33° S, and the Joint Authority Southern Demersal Gillnet and Demersal Longline Managed Fishery (JASDGDLF), which operates from 33° S to the WA/SA border. Sandbar shark and to a lower extent dusky shark were also targeted by the Northern Shark Fisheries, which operated between 114° 06' E and the WA/Northern Territory (NT) border and ceased operations in 2008-09. The current catch of sharks and rays in other WA commercial fisheries is negligible (< 10 t per annum) and recreational fishers retain very small numbers of sharks in WA. Indigenous catches of these species are considered to have always been negligible.

Based on their inherent vulnerability and risk to the sustainability, whiskery, gummy, dusky and sandbar sharks have been selected as indicators for the status of the temperate elasmobranch (sharks and rays) ‘suite’ as they represent the range of life history strategies of other elasmobranch species caught by these fisheries. However, with the increasing number of elasmobranch species being listed in national (Environment Protection and Biodiversity Conservation Act 1999, EBPC Act) and international (Convention on International Trade in Endangered Species, CITES; Convention on the Conservation of Migratory Species, CMS) protection lists, species-specific scientific advice on stock status at the local level is needed to comply with national and international legislation.  

Following a comprehensive reconstruction of total catch trajectories, a range of assessment methods was used to provide species-specific scientific advice on the stock status of all elasmobranch species identified in the reconstructed catches. First, a Productivity Susceptibility Analysis (PSA) was used to filter out species with historically low catch volumes from further analysis. Second, species ranked by the PSA as having high vulnerability were assessed using a combination of spatio-temporal changes in reported catch and effort, catch-only models, State-space surplus production models, and age- and sex-structured integrated models, depending on data availability (Figure 1). Biomass estimates from the production and integrated models were compared to reference levels relating to estimates of Maximum Sustainable Yield (MSY). The estimated biomass expected to achieve MSY (B~MSY~) is considered as the Threshold reference level (B~Thr~) for the stock, 0.5 B~MSY~ is set as the Limit reference level (B~Lim~) and 1.2 B~MSY~ was set as the Target reference level (B~Tar~). A flowchart explaining the steps undertaken in this assessment is shown in Figure 2.

**REDO FROM HERE............** For whiskery and gummy sharks, an integrated size-based model was implemented in 2017 to extend previous assessment models. The integrated model incorporated life history, gear selectivity, size composition, growth, catch and standardised fishery-dependent catch rate data up to and including the 2015-16 financial year. In addition, life history and catch information were used in a combined demographic and stock-reduction model (SR) to assess stock sustainability. Standardised fishery-dependent catch rate of gummy sharks was concluded to be a poor index of population abundance. For dusky and sandbar sharks, the time series of catch and effort data were considered to be insufficient for estimating biomass trends from fitting population dynamics models. Hence, a SR modelling approach was implemented to determine catch sustainability using life history and catch information up to and including the 2015-16 financial year.

The integrated model estimate for fishing mortality in 2021 was 0.024 year-1, with the 95% CLs ranging from 0.011 to 0.037 year-1. As the current value of this performance indicator is well below the estimated level of F corresponding to B~MSY~ (i.e., F~MSY~), the stock is unlikely to deplete to a level at which recruitment could be impaired if the current historically low level of catch is maintained. The integrated model estimate for female spawning biomass in 2021 was at 0.27 of the unfished level (95% CLs = 0.13-0.42). Although the current value of this performance indicator is below the threshold of 0.3, Gascoyne Oceanic Snapper is expected to recover to above this level over the next five years if the current low level of catch is maintained.

Based on all available lines of evidence (see Table below), the risk level for Gascoyne Oceanic Snapper over the next five years is estimated to be **Medium**.
On the basis of the evidence, Gascoyne Oceanic Snapper is classified as a  **Recovering stock**.**............. TO HERE**


For example, for grey nurse shark and shortfin mako, the inferred K equates to a virgin population size of ~7,000 individuals each across WA. 

```{r,echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.RAR,'/Spatial_LoE_Table.csv',sep=''))

flextable(Tab)%>%
      set_header_labels(Blocks.fished="Blocks\nfished",
                        Blocks.fished.with.catch="Blocks fished\nwith catch")%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(j='Effort',width=2,unit='cm')%>%
  width(j='Blocks.fished',width=3,unit='cm')%>%
  width(j='Blocks.fished.with.catch',width=3,unit='cm')%>%
  width(j='Consequence',width=3,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=c('Effort'),align='left',part='all')%>%
  set_caption(caption = 'Table A.6.1. Derived Consequence and Likelihood levels from the spatial distribution of catch and effort (NB for species for which market value decreased or they became protected, ‘Blocks fished with catch’ were set to ‘Stable’).',
              autonum = NULL)
```

<br>

Finally, risk scores were then calculated as the product of each consequence level and a weighted mean likelihood. Spatio-temporal changes in catch and effort, COMs, state-space surplus production models and integrated population models were given a weight of 0.25, 0.5, 1 and 2, respectively. The highest score was chosen as the overall fishing sustainability risk following the precautionary approach.


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Available time series by species.tiff',sep='')

```
![Figure 1. Available time series for the species assessed as High vulnerability by the PSA. F, empirical fishing mortality estimates for the TDGDLF.](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.an,'/Stock assessment flow chart.tiff',sep='')

```
![Figure 2. Flowchart describing the steps undertaken in the current assessment.](`r path`){width=800px}



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Catch-MSY
The F/F~MSY~ trajectories for indicator species are shown in Figure A.10.5 and for non-indicator species are shown in Figure A.10.6.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.over.Fmsy_catch.only_Indicator.sp.tiff',sep='')
```
![Figure A.10.5. Time series of F over FMSY (60% and 95% confidence intervals shown in dark grey and light grey, respectively) inferred from the catch-only model ensemble for indicator species. Also shown is the threshold reference point.](`r path`){width=800px}

<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.over.Fmsy_catch.only_Other.sp.tiff',sep='')
```
![Figure A.10.6. Time series of F over FMSY (60% and 95% confidence intervals shown in dark grey and light grey, respectively) inferred from the catch-only model ensemble for non-indicator species. Also shown is the threshold reference point.](`r path`){width=800px}


The B/BMSY trajectories inferred from the COM ensemble showed considerable uncertainty mostly due to the wide Bf/K range used to minimize constraining the models. The B/BMSY trajectories started to decline mostly around the late 1970s and early 1980s following initial exploitation and then increased or stabilised following the reduction in catches in recent years (Figure A.10.4). For 2021-22, inferred median B/BMSY was at or above the target reference level for most species, above threshold levels for spurdogs and at threshold levels for angel, copper and pigeye sharks.    

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/B.over.Bmsy_catch.only_Indicator.sp.tiff',sep='')
```
![Figure A.10.3. Time series of B over BMSY (60% and 95% confidence intervals shown in dark grey and light grey, respectively) inferred from the catch-only model ensemble for indicator species. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/B.over.Bmsy_catch.only_Other.sp.tiff',sep='')
```
![Figure A.10.4. Time series of B over BMSY (60% and 95% confidence intervals shown in dark grey and light grey, respectively) inferred from the catch-only model ensemble for non-indicator species. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}


These scenarios were compared on the basis of Deviance Information Criterion (DIC) and Root-Mean-Squared-Error (RMSE) values and overall biomass trajectories.

A comparison of the goodness of fit statistics for the different sensitivity tests done is shown in Figure A.11.1 for indicator and Figure A.11.2 for non-indicator species. For all species, the different scenarios considered showed very similar DIC and RMSE values with the exception of S4 (lower process error), which shower higher values.

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/GoodnessFit per scenario_JABBA CPUE_Indicator.sp.tiff',sep='')
```
![Figure A.11.1. Goodness of fit statistics for the different sensitivity tests conducted for indicator species. S2, lower r value; S3, higher r value; S4, smaller process error; S5, uniform K prior; S6, Schaefer production function; S7; Fox production function. DIC, Deviance Information Criterion; RMSE, Root Mean Squared Error.](`r path`){width=600px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/GoodnessFit per scenario_JABBA CPUE_Other.sp.tiff',sep='')
```
![Figure A.11.2. Goodness of fit statistics for the different sensitivity tests conducted for non-indicator species. S2, lower r value; S3, higher r value; S4, smaller process error; S5, uniform K prior; S6, Schaefer production function; S7; Fox production function. DIC, Deviance Information Criterion; RMSE, Root Mean Squared Error.](`r path`){width=600px}

<br>

<!-- Display Kobe plots for S1--> 
For the base case scenario, Kobe plots for indicator species are shown in Figure A.11.16.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_JABBA CPUE_Indicator.sp.tiff',sep='')
```
![Figure A.11.16. Kobe plots indicating the estimated trajectories of B over BMSY and F over FMSY for indicator species. The legends show the credibility intervals for the terminal assessment year (grey areas) and the probability of the terminal year falling within each quadrant.](`r path`){width=800px}

For the base case scenario, Kobe plots for non-indicator species are shown in Figure A.11.17.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_JABBA CPUE_Other.sp.tiff',sep='')
```
![Figure A.11.17. Kobe plots indicating the estimated trajectories of B over BMSY and F over FMSY for non-indicator species. The legends show the credibility intervals for the terminal assessment year (grey areas) and the probability of the terminal year falling within each quadrant.](`r path`){width=800px}

<br>

For wobbegongs, as only large individuals are retained (@bracciniQuantifyingSharkRay2021), a retention function was also implemented with discard mortality pre-specified at 10% (@bracciniHighPostcaptureSurvival2012).
For copper shark, length composition was also available for South Australia Marine Scalefish Fishery, the main source of fishing mortality for this species.


For sandbar sharks, only the survey series was used as this reflects breeding biomass trends whereas TDGLDF catch rates reflect small juveniles' biomass due to strong gillnet selectivity effects (@Simpfendorfer1998c, @McAuley2007a).




For the base case scenario, there was an overall negligible retrospective pattern, producing consistent stock status estimates (Figure A.11.8) and hence supporting the use of the base case model for making future projections. However, for dusky and spinner sharks the Mohn's-rho (ρ) statistic fell outside the expected interval for long-lived species of -0.15 and 0.2 (@carvalhoCookbookUsingModel2021) for B/B~MSY~ and B, respectively, indicating a degree of retrospective pattern although retrospective peels fell within the estimated 95% CI from the reference run, which is consistent with the uncertainty introduced by removing years of data

For the base case scenario, the hindcasting cross-validation results showed predictions fell within the 95% CI suggesting good predicting skills overall (Figure A.11.9).  For copper, gummy and spinner sharks, the mean absolute scale error (MASE) score was > 1 (the reference level), which implies that  the average model forecast is no better than a random walk (@carvalhoCookbookUsingModel2021). However, the predictions fell within the 95% CI and that for both species, the indices have a flat trend. For the other species, the MASE scores were ≤ 1 (i.e., predictions are better than a random walk) although for dusky shark most predicted years fell outside the 95% CI.


The alternative M scenarios used in the demographic analysis were used as a sensitivity (Table A.10.1).
Two *r* scenarios, resulting from the alternative M scenarios used in the demographic analysis, were used as a sensitivity (Table A.10.2).



```{r, fig.width = 5, fig.height=5, fig.cap = "Figure A.4.1. Proportional shark and ray catch for Western Australia fisheries. The top panel shows the proportion by method for shark and non-shark fisheries. The bottom panel is a word cloud of the different fisheries interacting with sharks and rays. The southern and northern shark fisheries are shown in green and orange, respectively, whereas non-shark fisheries are shown in black. The size of the word is proportional to the square root of the total catch.", fig.pos='h',fig.align='center'}
knitr::include_graphics(paste(Model.outputs.an,'Proportional catch_WA_combined.tiff',sep=''))
```


```{r, fig.width = 7, fig.height=8, fig.cap = "Figure A.5. Total reported effort for the TDGDLF (top pane) and NSF (bottom pane). Also shown are the management measures and data collection initiatives implemented since the start of these fisheries. Effort is shown in a logarithmic scale. Red, data collection; Blue, spatial closures; Green, other management measures.", fig.pos='h',fig.align='center'}
knitr::include_graphics(paste(handl_OneDrive('Management/Sharks/Timeline management measures/'),'Effort.management_TDGDLF & NSF.tiff',sep=''))
```
