---
title: "MEMORANDUM"
subtitle: '(not for distribution outside DPIRD)'
output:
  officedown::rdocx_document  
bibliography: 'C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Analyses/Build-Documents/My Library.bib'
csl: 'C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Analyses/Build-Documents/canadian-journal-of-fisheries-and-aquatic-sciences.csl'
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Reports/WoE Advice/2023") })
header-includes:
  - \pagenumbering{gobble}

---

```{r globaloptions, include=FALSE}
this.yr=2022      # assessment year
report.yr=2023    # report year
source('C:/Users/myb/OneDrive - Department of Primary Industries and Regional Development/Matias/Analyses/SOURCE_SCRIPTS/Git_other/handl_OneDrive.R')
library(knitr)
library(tidyverse)
library(readxl)
library(Hmisc)
library(officedown)
library(flextable)
library(stringi)
library(Hmisc)
library(officer)
library(r4ss)
# https://www.zotero.org/styles?fields=biology  Zotero Citation Style Language
```

<!-- global for setting path to figures and formattig -->
```{r, echo=FALSE}
#outputs in RAR folder
Model.outputs.RAR=handl_OneDrive(paste('Reports/RARs',this.yr,'Figures_and_Tables',sep='/'))   
Model.risk.tabs.RAR=handl_OneDrive(paste('Reports/RARs',this.yr,'Risk tables',sep='/'))

#outputs in Analysis folder
Model.outputs.an=handl_OneDrive('Analyses/Population dynamics/')       

#location of static figures/tables (i.e. background info)
Repository=handl_OneDrive('Reports/RARs/Repository')   

Indicator.species=c("Gummy shark","Dusky shark","Whiskery shark","Sandbar shark")

year_current_future<- read.csv(paste(Model.outputs.RAR,'/year_current_future.csv',sep=''))
year.current=year_current_future$year.current
year.future=year_current_future$year.future
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, ft.keepnext = F)  #chunk for keep table next to caption
knitr::opts_chunk$set(tab.cap.pre = "", tab.cap.sep = "")   #remove Table prefix
knitr::opts_chunk$set(comment = '', fig.width = 10,fig.height = 10)   #figure dimensions
knitr::opts_chunk$set(number_sections = FALSE)  #suppressing numbering of subheadings
```

\pagenumbering{arabic}
To	      Executive Director – Fisheries and Agriculture Resource Assessment
	
From	    Chief Fisheries Scientist – Aquatic Science and Assessment
	
`r paste('Subject ELASMOBRANCH RESOURCE -' ,format(Sys.time(), "%Y"), 'ASSESSMENT')` 	  
	
File No	xx-xxxxx
	
CC	SPRS Offshore; PRS Demersal; SPRS SADA
	
Date	`r format(Sys.time(), "%d %B %Y")` 


# Background
This document provides the `r format(Sys.time(), "%Y")` research advice on the stock status and risk to sustainability over the next five years for all elasmobranch (sharks and rays) species and/or stocks identified in the catches of commercial and/or recreational fisheries operating in Western Australia (WA). This includes the indicator species, gummy (*Mustelus antarcticus*), dusky (*Carcharhinus obscurus*), whiskery (*Furgaleus macki*) and sandbar (*C. plumbeus*) sharks, and non-indicator species, i.e., all other elasmobranch species caught. A risk-based Weight of Evidence (WoE) assessment of stocks status was undertaken for all these species using ISO 31000-based risk assessment methods.

# Current situation
## Indicator Species
The current assessment for gummy shark is based on catch-only modelling and spatio-temporal catch and effort distribution (Level 1), state space surplus production modelling fitted to the abundance series (Level 4) and integrated population dynamics modelling (Level 5) fitted to the abundance series, length composition and mean body weight data. Based on all available lines of evidence (see Weight of Evidence Tables below), the risk level for gummy shark over the next five years is estimated to be **Low**. Current (2021-22) biomass estimates exhibit considerable uncertainty, but estimates are above the target level and are projected to increase. Current fishing mortality estimates are below the target level and are projected to remain below this level. 

On the basis of this evidence, gummy shark is classified as a **Sustainable stock**. 

The current assessment for dusky shark is based on catch-only modelling and spatio-temporal catch and effort distribution (Level 1), and integrated population dynamics modelling (Level 5) fitted to length composition data. Based on all available lines of evidence (see Weight of Evidence Tables below), the risk level for dusky shark over the next five years is estimated to be **High**. Current (2021-22) median biomass is at the threshold level and projected to increase whereas median fishing mortality is below the target level and projected to remain slightly below this level. However, model estimates showed considerable uncertainty and the integrated model was fitted to length composition data only as the available abundance series was considered unreliable for this stock. 

On the basis of this evidence, dusky shark is classified as a **Recovering stock**. 

The current assessment for whiskery shark is based on catch-only modelling and spatio-temporal catch and effort distribution (Level 1), state space surplus production modelling fitted to the abundance series (Level 4) and integrated population dynamics modelling (Level 5) fitted to the abundance series, length composition and mean body weight data. Based on all available lines of evidence (see Weight of Evidence Tables below), the risk level for whiskery shark over the next five years is estimated to be **Medium**. Current (2021-22) biomass estimates showed considerable uncertainty but median estimates are at the target level and are projected to increase. Current fishing mortality estimates are below the target level and are projected to decrease. 

On the basis of this evidence, whiskery shark is classified as a **Sustainable stock**. 

The current assessment for sandbar shark is based on catch-only modelling and spatio-temporal catch and effort distribution (Level 1), state space surplus production modelling fitted to the abundance series (Level 4) and integrated population dynamics modelling (Level 5) fitted to the abundance series, length composition and mean body weight data. Based on all available lines of evidence (see Weight of Evidence Tables below), the risk level for sandbar shark over the next five years is estimated to be **Low**. Current (2021-22) biomass estimates are above the target level and are projected to increase. Current fishing mortality estimates are below the target level and are projected to remain below this level. 

On the basis of this evidence, sandbar shark is classified as a **Sustainable stock**.

For the indicator species, maintaining current management settings (i.e., assuming total catches are sustained at current levels) is predicted to maintain the sustainability risk at acceptable levels. However, biomass trajectories show considerable uncertainty due to the data-limited nature of the assessments and this should be a consideration for future management arrangements.

## Non-indicator Species
The current assessment for non-indicator species is based on catch-only modelling and spatio-temporal catch and effort distribution (Level 1), state space surplus production modelling fitted to the abundance series (Level 4) and/or integrated population dynamics modelling (Level 5) fitted to the abundance series and/or length composition data depending on data availability. Based on all available lines of evidence (see Weight of Evidence Tables below), the risk level over the next five years is estimated to be **High** for angel sharks, **Medium** for sawsharks, copper, and spinner sharks, and  **Low** for the other non-indicator species. Non-indicator species are only a minor component of fisheries currently operating under the jurisdiction of WA. The highest sources of fishing mortality for these species ceased decades ago. For two of the non-indicator species (copper and angel sharks), most of the catch has been taken by fisheries outside the jurisdiction of WA. Copper shark is a shared stock across southern Australia and most of the catch originates from South Australia’s Marine Scalefish Fishery. Angel sharks in WA are mainly captured by the Commonwealth managed Great Australian Bight Trawl Fishery and were classed as **High** risk due to the high levels of uncertainty in the MSY value inferred from catch-only modelling (see *Level 1 Assessment* below in *Risk-based weight of evidence table: Non-indicator species*). 

On the basis of this evidence, all non-indicator species are classified as **Sustainable stocks**.

\newpage

# Risk-based Weight of Evidence Assessment of the Elasmobranch Resources of Western Australia

Assessment date: `r format(Sys.time(), "%B %Y")`

# Executive Summary
Over 100 elasmobranch species have been identified in the catches of commercial and/or recreational fisheries in Western Australia (WA). Gummy, dusky, whiskery, and sandbar sharks are the most commonly captured species (~80% of the elasmobranch catch). These species have been targeted in the Temperate Demersal Gillnet and Demersal Longline Fisheries (TDGDLF), which operates in the West Coast and South Coast Bioregions and comprises the West Coast Demersal Gillnet and Demersal Longline (Interim) Managed Fishery (WCDGDLF), which operates between 26° and 33° S, and the Southern Demersal Gillnet and Demersal Longline Managed Fishery (SDGDLF), which operates from 33° S to the WA/South Australia (SA) border. Sandbar shark and, to a lesser extent, dusky shark were also targeted by the Northern Shark Fisheries, which operated between 114° 06’ E and the WA/Northern Territory (NT) border and ceased operations in 2008-09. The current catch of sharks and rays in other WA commercial fisheries is negligible (< 10 t per annum), and recreational fishers retain very small numbers of sharks in WA. Indigenous catches of these species are considered to be negligible both in terms of historical and extant harvest.

Based on their importance to fisheries, inherent vulnerabilities, and sustainability risk, whiskery, gummy, dusky, and sandbar sharks have been selected as indicators for the status of the temperate elasmobranch (sharks and rays) ‘suite’ as they represent the range of life history strategies of other elasmobranch species caught by these fisheries. However, with the increasing number of elasmobranch species being listed in national (Environment Protection and Biodiversity Conservation Act 1999, EBPC Act) and international (Convention on International Trade in Endangered Species, CITES; Convention on the Conservation of Migratory Species, CMS) protection lists, species-specific scientific advice on stock status at the local level is needed to understand species status in the context of national and international species conservation significance.

For whiskery shark, the initial historical stock assessment [@Simpfendorfer2000a] was based on a bespoke age- and sex-structured model (i.e., L5 assessment). This assessment determined that the mature female biomass in 1997-98 (the last year of the assessed data series) was at 23% of virgin levels, and to achieve management biomass targets with a 50% probability by 2010, recommended catches were required to be maintained below 190 tonnes. For sandbar and dusky sharks, previous stock assessments [@McAuley2007b] were based on demographic and mark recapture analyses. These assessments determined that the combined fishing pressure from WA shark fisheries and cryptic mortality from other fisheries was deemed unsustainable at the effort peak (1990s-2000s). More recently, the weight of evidence assessments for the four indicator species (L1 and L5 assessments for whiskery shark, L1 assessments for gummy, dusky, and sandbar sharks; @bracciniRiskbasedWeightEvidence2021), determined that the risk posed by fishing to the sustainability of these stocks was assessed as Medium for 2015-16 (the last year of the data series) and for the projected period to 2020-21. Prior to the current study, there has been no stock assessments for other shark and ray species captured in WA because the main species were considered as indicators to entire suite of species.

Following a comprehensive reconstruction of the total catch trajectories of all species for which data are available, a range of assessment methods was used to provide species-specific scientific advice on the stock status of all elasmobranch species in WA. First, a Productivity Susceptibility Analysis (PSA) was used to filter out species with historically very low catch volumes, and those with low vulnerability from further analysis. Second, species ranked by the PSA as having high vulnerability were assessed using a combination of spatio-temporal changes in reported catch and effort, catch-only models, State-space surplus production models, and age- and sex-structured integrated models, depending on data availability (Figure 1). Biomass estimates from the production and integrated models were compared to reference levels relating to estimates of Maximum Sustainable Yield (MSY). The estimated biomass expected to achieve MSY (BMSY) was considered as the Threshold reference level (BThr) for the stock; 0.5 BMSY was set as the Limit reference level (BLim) and 1.2 BMSY was set as the Target reference level (BTar). A flowchart explaining the steps undertaken in this assessment is shown in Figure 2.

For the indicator species, the major performance measures for assessing stock status are spawning biomass levels and fishing mortality (F), estimated using an integrated age-structured assessment model. For the non-indicator species, a range of performance measures for assessing stock status are used based on data availability.

For gummy shark, the estimated biomass and F trajectories showed considerable uncertainty. Biomass trajectories declined considerably in the 1980s following increasing exploitation but the median biomass estimate remained above the target level throughout the time series. Projections to 2026-27 show a slight increasing trend in levels of biomass. For 2021-22 and 2026-27, the probability that the relative biomass was above the limit reference level was 100%, and the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Low**. On the basis of this evidence, gummy shark is classified as a **Sustainable stock**.

For dusky shark, biomass trajectories declined considerably in the late 1970s following increasing exploitation, with the median biomass estimate falling below the threshold reference level in the 2010s and increasing in recent years following a reduction in catches, with the median biomass estimate reaching threshold levels in 2020-21. Projections to 2026-27 show a slight increasing trend in levels of biomass. For 2021-22 and 2026-27, the probability that relative biomass was above the limit reference level was 100%. However, the probability that relative biomass was between the limit and threshold reference levels was 47% for 2021-22, and 43% for 2026-27. The F trajectories remained at a negligible level at the start of the time series, peaking in the late 1980s and 1990s (when limit reference levels were exceeded), and declining in recent years. For 2021-22, the median F was just below the target reference level. Projections to 2026-27 showed a stable trend in F estimates. For 2021-22 and 2026-27, the probability that F was below the limit reference level was 100%. However, the probability that F was between the threshold and target levels was 31% for 2026-27. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **High**, mostly due to estimation uncertainty as the integrated model was fitted to length composition data only. On the basis of this evidence, dusky shark is classified as a **Recovering stock**.

For whiskery shark, biomass trajectories declined considerably in the late 1940s following increasing exploitation, with the median biomass estimate falling below the limit reference level between the 1990s and early 2010s, before increasing in recent years following a reduction in catches, with the median biomass estimate remaining above target reference levels since 2020-21. Projections to 2026-27 show an increasing biomass trend. For 2021-22 and 2026-27, the probability that relative biomass was above the limit reference level was 87% and 97%, respectively. However, the probability that relative biomass was below the limit, and between the limit and threshold reference levels was 13% and 19%, respectively, for 2021-22. The F trajectories remained at a negligible level at the start of the time series, peaking in the 1990s (when limit reference levels were exceeded), and declining in recent years. For 2021-22, median F was considerably below the target reference level. Projections to 2026-27 show a declining trend in F (well below the target reference level). For 2021-22 and 2026-27, the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Medium**. On the basis of this evidence, whiskery shark is classified as a **Sustainable stock**.

For sandbar shark, biomass trajectories declined considerably in the 1970s following the initial period of exploitation, with the median biomass estimate falling between the threshold and limit reference levels in the late-2000s. However, median biomass increased in recent years following a reduction in catches,remaining above target reference levels since the early 2010s. Projections to 2026-27 show an increasing biomass trend. The F trajectories remained at a negligible level at the start of the time series, peaking in the late 2000s (when limit reference levels were exceeded) and declining in recent years. For 2021-22, median F was considerably below the target reference level. Projections to 2026-27 show a stable trend in levels of F (i.e., below the target reference level). For 2021-22 and 2026-27, the probability that the relative biomass was above the limit reference level was 100%, and the probability that F was below the limit reference level was 100%. Based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be **Low**. On the basis of this evidence, sandbar shark is classified as a **Sustainable stock**.

For the non-indicator species, based on all available lines of evidence (see Table below), the risk level over the next five years is estimated to be High for angel sharks, **Medium** for sawsharks, copper, and spinner sharks, and Low for the other non-indicator species. For copper and angel sharks, most of the catch has been taken by fisheries outside the jurisdiction and management influence of WA, and angel sharks were classed as **High** risk due to the high levels of uncertainty in the MSY value inferred from catch-only modelling (see Level 1 Assessment below in Risk-based weight of evidence table: Non-indicator species). On the basis of this evidence, all non-indicator species are classified as **Sustainable stocks**.



```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Available time series by species.tiff',sep='')

```
![Figure 1. Available time series for the species assessed as having High vulnerability to fishing by the PSA. F, empirical fishing mortality estimates for the TDGDLF. (N.B., F was obtained from McAuley et al., 2007a)](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.an,'/Stock assessment flow chart.tiff',sep='')

```
![Figure 2. Flowchart describing the steps undertaken in the current assessment. The number of species assessed by the different assessment methods is shown in brackets. (N.B., PSA vulnerability thresholds for defining risk level were obtained from Hobday et al. (2007))](`r path`){width=800px}


# Future Monitoring & Assessment
Additional information (e.g., a mortality estimate for a certain period from representative age composition data, a point estimate of absolute abundance from genetic techniques such as close-kin mark-recapture, Bravington et al. [-@bravingtonAbsoluteAbundanceSouthern2016]), would assist in reducing model uncertainties, particularly for those species for which reconstructed catches are not negligible but only catch and life history information are available. Also, life history information for most species is not stock-specific, and in a few cases, not species-specific. Hence, targeted collection of biological information (reproduction, growth) would assist future assessments as stock-specific information that is currently available is several decades old. This is particularly relevant for dusky shark, which was classified as High risk.

There is a heavy reliance on fishery-dependent catch per unit of effort (CPUE) data to inform depletion levels for all species except sandbar shark, for which reliable fishery-independent data are available. A long-term fishery-independent monitoring program to collect abundance and/or composition data, particularly in temperate WA, would assist in reducing model uncertainties. This would be particularly useful for gummy, whiskery, and dusky sharks, noting that some of the integrated model fit diagnostics were not robust (e.g., retrospective analysis and hindcasting for gummy shark, likelihood profiling for most assessed species) due to data limitations or could not be conducted, such as retrospective analysis and hindcasting, due to lack of recent data (e.g., dusky shark).

From a modelling viewpoint, future surplus production models could explore the use of JABBA-select to allow for the incorporation of CPUE series from fisheries with different selectivity schedules. It is also recommended that future integrated population models attempt to better account for the spatial structure of the TDGDLF and the use of two different gillnet mesh sizes (6.5 and 7 inch). Future integrated models could also attempt to estimate Fishing mortality from published mark-recaptured data rather than input it as an additional CPUE series (current approach).

Unaccounted illegal fishing in northern WA is a source of concern. The implemented catch reconstructions use the annual number of boat apprehensions to scale up observed shark catch volumes and species composition from a sample of apprehended vessels. In recent years, the number of boat apprehensions decreased to almost 0 due to change in policies but the number of legislative forfeitures increased to a historic high, from close to 0 in 2019 to >300 in 2021. Future assessments should explore the implications of using the number of legislative forfeitures in the calculation of reconstructed catches and use those updated figures in the estimation of populations trajectories.

For many of the non-indicator species, the MSY values inferred from the catch-only assessments are likely to be underestimated given the historically low levels of population extractions across the species distribution in WA. If specific management arrangements (e.g., setting acceptable catch ranges) are required for these non-indicator species, targeted research effort (e.g. mortality estimates for representative age composition data) should be implemented. 

Finally, a formal harvest strategy that considers and defines reference points and control rules for managing the resource is required, as the current management settings only provide a single reference point for assessing shark species (40% unfished biomass), and this reference point is not explicitly defined as a target, threshold, or limit. In addition, unlike other resources for which reference points relate to the spawning biomass (equivalent to female mature biomass for sharks), the current reference point in the TDGDLF relates to total biomass. A successful harvest strategy could also be informed through Management Strategy Evaluation (e.g., Doering et al. [-@doeringSSMSEPackageManagement2023]).  

The next assessment for the WA Elasmobranch Resource is proposed for 2029 to allow for data collection, processing and analyses.


# Assessed species
Including the four indicator species, >100 shark and ray taxonomic groups (herein referred to as 'species') have been identified in the catch of recreational and/or commercial fisheries operating in WA (Table A.2). However, only a subset of those are currently assessed because quantitative population assessment approaches are not suitable for species with negligible catches [@Martell2012] and some species were assessed elsewhere. Blacktip (*C. limbatus* and *C. tilstoni*) and spot-tail (*C. sorrah*) sharks form a single biological stock between WA and NT, and WA catches are a small fraction of the total catch; hence, these species are assessed as part of the assessment conducted by the Fisheries Division, Department of Industry, Tourism and Trade, Northern Territory [@Stewardson2018]. Similarly, school shark (*Galeorhinus galeus*) form a single biological stock throughout southern Australia and the WA catches are negligible compared to other jurisdictions. Hence, this species is assessed by the Commonwealth of Australia [@Stewardson2018]. Next, white shark (*Carcharodon carcharias*), which were reported as landed catch in commercial logbooks until their protection in 1997, were assessed by Hillary et al. [-@hillaryGeneticRelatednessReveals2018]. For bull shark (*C. leucas*), reported catches (mostly taken in the NSF), are negligible and unreliable as this species was not observed during three years of intensive on-board observer sampling or during several years of fishery-independent shark longline surveys (Heupel and McAuley, 2007). Hence, bull shark catches were combined with those of pigeye shark (*C. amboinensis*), and assessed as ‘pigeye’, given morphological similarities between these species [@tillettSimilarLifeHistory2011] and the validated presence of pigeye sharks in the catches of the northern shark fisheries [@Heupel2007b]. Finally, due to variable resolution in species identification across the different catch sources, catsharks (Scyliorhinidae), sawsharks (Pristiophoridae), thresher sharks (Alopidae), wobbegongs (Orectolobidae), angel sharks (Squatinidae), banjo rays (Trygonorrhinidae), and wedgefish (Rhinidae) were assessed at the family level. All analyses were undertaken using the statistical software R [@RDevelopmentCoreTeam2014], using a range of packages described in the corresponding Appendices. 


```{r, echo=FALSE}
assessed.species<- read.csv(paste(Model.outputs.RAR,'/Assessed_species.csv',sep=''))
assessed.species<-assessed.species%>%pull(Species)
names.assessed.species<- paste(paste(assessed.species[1:(length(assessed.species)-1)],collapse=', '),
                             'and',assessed.species[length(assessed.species)])
```

A PSA assessment was undertaken to remove species with historically low catches and/or those species with either low or medium vulnerability from quantitative assessments (Appendix 3). Of the >100 species considered in this assessment, the following species with a high vulnerability to fishing were included for subsequent assessment: `r tolower(names.assessed.species)` (Figure 3).   


```{r, echo=FALSE}
path=paste(Model.outputs.an,'PSA/Figure_PSA.tiff',sep='')

```
![Figure 3. PSA for the assessed species. Also shown is the resulting vulnerability threshold. The size of the species name label is proportional to the vulnerability score.](`r path`){width=800px}

<br>


# Risk-Based Weight of Evidence Tables
<!-- Populate each species' Weight of evidence table and Risk-based weight of evidence -->
```{r, echo=FALSE, results ='asis', ft.align="left"}
get.excel.tabs=function(PATH)
{
  tab_names <- excel_sheets(path = PATH)
  data.list=lapply(tab_names, function(x) read_excel(path = PATH, sheet = x))
  names(data.list)=tab_names
  return(data.list)
}

WoE.tables=get.excel.tabs(paste(Model.risk.tabs.RAR,'1.Weight of evidence tables.xlsx',sep='/'))

Species=names(WoE.tables)
for(i in 1:length(Species))
{
  cat('\n##', Species[i], '\n')
  #cat('\n##', paste('Risk-based weight of evidence table:',Species[i]), '\n')
  flextable(WoE.tables[[i]])%>%
  width(j='Category',width=3,unit='cm')%>%
  width(j='Lines of evidence',width=14,unit='cm')%>%
  align(align="left")%>%
  valign(valign = "top", part = "all")%>%
  hline(part="all")%>%
  vline(part="all")%>%
  border_outer(part="all")%>%
  merge_at(i = 3, j = 1:2 )%>%
  merge_at(i = 4, j = 1:2 )%>%
  merge_at(i = 7, j = 1:2 )%>%
  merge_at(i = 8, j = 1:2 )%>%
  merge_at(i = 9, j = 1:2 )%>%
  merge_at(i = 12, j = 1:2 )%>%
  merge_at(i = 13, j = 1:2 )%>%
  merge_at(i = 15, j = 1:2 )%>%
  merge_at(i = 16, j = 1:2 )%>%
  merge_at(i = 19, j = 1:2 )%>%
  merge_at(i = 20, j = 1:2 )%>%
  merge_at(i = 21, j = 1:2 )%>%
  merge_at(i = 22, j = 1:2 )%>%
  merge_at(i = 23, j = 1:2 )%>%
  merge_at(i = 24, j = 1:2 )%>%
  merge_at(i = 25, j = 1:2 )%>%
  merge_at(i = 26, j = 1:2 )%>%
  bold(i=c(3,7,9,12,15,19,21,22,24,25))%>%
  bg(i=c(3:4,7:9,12:13,15:16,19:24),bg='grey85')%>%
  bg(i=c(25:26),bg='lightblue2')%>%
  bg(bg = 'grey85', part = "header")%>%
  bold(part = "header")%>%
  bold(i=c(1:2,5:6,10:11,14,17:18),j=1)%>%
  hline(i=c(3,7:8,15,17,19:20,22:23,25),border=fp_border(color="transparent"))%>%
    flextable_to_rmd

  cat('\n')
}

```

<br>


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Risk_Indicator.sp.tiff',sep='')

```
![Figure 4. Weight of evidence assessment for indicator species. BCur, Current biomass; BLim, limit biomass; BThr, threshold biomass; BTar, target biomass.](`r path`){width=800px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Risk_Other.sp.tiff',sep='')

```
![Figure 5. Weight of evidence assessment for non-indicator species. BCur, Current biomass; BLim, limit biomass; BThr, threshold biomass; BTar, target biomass.](`r path`){width=800px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Risk_all LoE for each species.tiff',sep='')

```
![Figure 6. Risk score for each line of evidence (LoE) and overall risk score for all species assessed. Species names were color coded: Bold brown, indicator species; Grey, non-indicator species; Black, species assessed only by the PSA. PSA, Productivity Susceptibility Analysis; Spatial; Spatio-temporal changes in catch and effort; JABBA, State Space Surplus Production Model; Integrated, Integrated Population Dynamics Model.](`r path`){width=800px}  

<br>

<!-- Display Final risk for each LoE and Overall--> 
The final risk for each line of evidence and the overall risk score is shown in Table 1 for each species.  
  
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 13. Risk of each LoE and Overall.csv',sep=''))
PSA.species=Tab1%>%filter(!is.na(PSA))%>%pull(Species)
Tab1.agg=Tab1%>%filter(Species%in%PSA.species)
Tab1.agg=Tab1.agg%>%
            distinct(Risk.overall,.keep_all = T)%>%
            mutate(Species=paste0('PSA-only species \n', '(n=',length(PSA.species),')'))
Tab1=Tab1%>%filter(!Species%in%PSA.species)%>%arrange(Species)
Tab=rbind(Tab1,Tab1.agg)
names(Tab)[match('Risk.overall',names(Tab))]='Overall risk'  
  
Tab%>% 
  flextable()%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=2,unit='cm')%>%
  width(j='Species',width=4,unit='cm')%>%
  set_caption(caption = "Table 1. Final risk score for each line of evidence and overall risk. PSA, Productivity Susceptibility Analysis; Spatial; Spatio-temporal changes in catch and effort; JABBA, State Space Surplus Production Model; Integrated, Integrated Population Dynamics Model.",
              autonum = NULL)

```
<br>

# References

<div id="refs"></div>

<br>

# Appendix 1. Consequence, Likelihood and Risk Levels (based on AS 4360 / ISO 31000) 
Following Fletcher [-@Fletcher2015], fishing sustainability risk was defined as negligible, low, medium, high or severe. These risk levels were calculated as the product of consequence and likelihood (see definition below). The likelihood of each consequence level was calculated for each line of evidence as detailed below. 

For spatio-temporal changes in catch and effort, likelihood values were derived from the spatial and temporal dynamics of the total number of blocks fished within each species' core area, from the proportion of blocks with catch, and from market dynamics and legislated species protection (Appendix 6). 

For data-limited catch only models (COMs), likelihood values were calculated based on probabilities of the total catch (for the last five years) of a species being above, between or below catch reference points. Rather than using a point estimate of total catch, a normal distribution was used with mean and standard deviation set at the mean and standard deviation of the observed total catch (for the last five years). For each species, the limit reference point for catch was set to the median value for MSY produced by the COMs, i.e., the limit reference point is considered to be breached if there is greater than 50% probability of MSY being exceeded. The threshold reference point was set to 0.8 of the MSY median value produced by the COMs, i.e., the threshold is considered to be breached if there is greater than 20% probability (and thus still appreciable risk) of MSY being exceeded. These limit and threshold catch reference points were derived based on the theoretical expectation that, for any stock, catches above MSY, over any period other than the short term, will result in overfishing and should thus be avoided [@Martell2012]. The target reference point was set to 0.5 of the MSY median value produced by the COMs, i.e., to a level of catch that is much lower than the median expectation for MSY and, taking into account full uncertainty for MSY from these data-limited approaches, was able to be shown for all species in this assessment to have zero chance of exceeding MSY.

For the state-space surplus production and integrated models, likelihood levels were derived from the model outputs. Risk scores were calculated for `r year.current` and `r year.future`. For this, the population models were projected for five years assuming a constant catch set at the average catch of the last five years. 

For each assessed species, risk scores for each consequence level were derived from the highest assessment level available interpreted within the context of the available lower assessment levels. The highest score was chosen as the overall fishing sustainability risk following the precautionary approach.



## Consequence levels
1. Minor – Fishing impacts either not detectable against background variability for this population; or if detectable, minimal impact on population size and none on dynamics. Biomass > Target level (B~MEY~)
1. Moderate – Fishery operating at maximum acceptable level of depletion. Biomass < Target level (B~MEY~) but > Threshold level (B~MSY~) 
1. High – Level of depletion unacceptable but still not affecting recruitment levels of stock. Biomass < Threshold level (B~MSY~) but >Limit level (B~REC~) 
1. Major – Level of depletion is already affecting (or will definitely affect) future recruitment potential/ levels of the stock. Biomass < Limit level (B~REC~) 

## Likelihood levels
These are defined as the likelihood of a particular consequence level actually occurring within the assessment period (5 years was used).

1. Remote – The consequence has never been heard of in these circumstances, but it is not impossible within the time frame (Probability of <5%)
2. Unlikely – The consequence is not expected to occur in the timeframe but it has been known to occur elsewhere under special circumstances (Probability of 5 - <20%)
3. Possible – Evidence to suggest this consequence level is possible and may occur in some circumstances within the timeframe. (Probability of 20 - <50%)
4. Likely – A particular consequence level is expected to occur in the timeframe (Probability of >50%)

```{r, echo=FALSE}
path=handl_OneDrive('Reports/RARs/Repository/Table_C_L.jpg')
```
![Figure A.1.1. Consequence - Likelihood table.](`r path`){width=800px}

<br>

```{r, echo=FALSE}
path=handl_OneDrive('Reports/RARs/Repository/Table_Risk.jpg')
```
![Figure A.1.2. Risk table.](`r path`){width=800px}

<br>


# Appendix 2. Assessed species
```{r,echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.an,'PSA/Table S1_All.species.caught.by.fishery.csv',sep=''))
Tab<- Tab%>%dplyr::select(Name,Scien.nm)%>%rename('Common name'=Name,'Scientific name'=Scien.nm)
flextable(Tab)%>%
  width(j='Common name',width=5,unit='cm')%>%
  width(j='Scientific name',width=6,unit='cm')%>%
  italic(j='Scientific name')%>%
  set_caption(caption = 'Table A.2. Shark and ray species identified in the catch of commercial and/or recreational fisheries operating in Western Australia.',
              autonum = NULL)
# knitr::kable(Tab, caption='Table A.2. Shark and ray taxonomic groups identified in the catch of commercial and/or recreational fisheries operating in Western Australia. Taxonomic groups are ordered alphabetically by common name.')

```
<br>


<!---BLOCK_LANDSCAPE_START--->
# Appendix 3. PSA scores
## Biology and vulnerability
Productivity scores (Table A.3.1) were determined using life history information on maturity, fecundity, reproductive strategy and trophic level (Table A.3.3), derived from life history information obtained from several scientific papers and reports (Appendix 9) and FishBase (https://www.fishbase.org).

```{r, echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.an,'PSA/Table S2_PSA productivity definition.csv',sep=''))  
names(Tab)=c("Productivity attribute","High productivity (Score= 1)","Medium productivity (Score= 2)",
"Low productivity (Score= 3)")
flextable(Tab)%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(j=1,width=4,unit='cm')%>%
  width(j=2:4,width=5,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=1,align='left',part='all')%>%
  valign(j=1:4, valign = "top", part = "all")%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.3.1. Productivity attributes.",autonum = NULL)

```

Susceptibility scores (Table A.3.2) were determined using information on species availability, encounterability, selectivity and post-capture mortality (PCM) in relation to the four main types of commercial fishing gears used in WA: gillnets, hooks, trawl and traps (Table A.3.3). The cumulative impact of these gears was accounted for by calculating an aggregated susceptibility index following Micheli et al. [-@Micheli2014]. Availability, encounterability and selectivity scores were derived from ecological risk assessments [@Walker2008; @Cortes2009; @Walker2021] and expert judgement whereas PCM scores were derived from PCM information published by Braccini et al. [-@bracciniHighPostcaptureSurvival2012] and Ellis et al. [-@Ellis2016]. 

```{r, echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.an,'PSA/Table S2_PSA susceptibility definition.csv',sep=''))
names(Tab)=c("Susceptibility attribute","Low susceptibility (Score= 1)","Medium susceptibility (Score= 2)","High susceptibility (Score= 3)")
flextable(Tab)%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(j=1,width=4,unit='cm')%>%
  width(j=2:4,width=5,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=1,align='left',part='all')%>%
  valign(j=1:4, valign = "top", part = "all")%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.3.2. Susceptibility attributes.",autonum = NULL)

```


Many species have been historically caught only sporadically and in very low volumes, indicating no sustainability risk [@Newman2018]. Hence, availability and encounterability scores were set to one for species with maximum annual catches < 50 tonnes and annual catches < five tonnes for ≥ five years. Annual catches from WA shark and ray species taken in WA fisheries were extracted from Braccini et al. [-@Braccini2021], Braccini & Murua [-@bracciniQuantifyingSharkRay2021] and Braccini et al. [-@bracciniRecreationalHarvestSharks2021]. These data include time series of reconstructed recreational catches, reported domestic commercial landings, unreported domestic discards, and illegal and unreported foreign catches within WA waters. For copper shark (*C. brachyurus*), catches from South Australia, New South Wales and Commonwealth Fisheries were also included in the assessment as this species forms a single unit-stock across southern Australia (Huveneers et al., 2021). 

Vulnerability scores were next calculated to determine which species should be included for further assessment. Vulnerability was calculated as the Euclidean distance from the origin to the coordinates of the productivity and susceptibility scores on the PSA plots [@cortesEXPANDEDECOLOGICALRISK2015]. Vulnerability thresholds for defining risk level were set at <2.64 (low risk), 2.64 to 3.18 (medium risk) and > 3.18 (high risk) [@Hobday2007]. Only species with ‘high’ vulnerability scores were considered in subsequent analyses. The indicator species were not considered in this analysis as their vulnerability score is high (a characteristic that defines their selection as an indicator species).

```{r,echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.an,'PSA/Table S2_PSA scores.csv',sep=''))
Nms=names(Tab)
flextable(Tab)%>%
  width(j='Species',width=3,unit='cm')%>%
  set_header_labels(
      Species="",
      Max.age="MA",Age.mat='AM',Fecun='FE',Max.size='MS',Size.mat='SM',Rep.strat='RS',Troph.Lvl='TL',
      Net.avail='Av',Net.encoun='En',Net.sel='Sel',Net.PCM='PCM',
      Line.avail='Av',Line.encoun='En',Line.sel='Sel',Line.PCM='PCM',
      Trawl.avail='Av',Trawl.encoun='En',Trawl.sel='Sel',Trawl.PCM='PCM',
      Trap.avail='Av',Trap.encoun='En',Trap.sel='Sel',Trap.PCM='PCM',
      Vulnerability.score="")%>%
  width(j=Nms[-match("Species",Nms)],width=.95,unit='cm')%>%
  width(j=Nms[grep('PCM',Nms)],width=1.25,unit='cm')%>%
  width(j="Vulnerability.score",width=1,unit='cm')%>%
  fontsize(size=9, part='all')%>%
  add_header_row(values = c("Species", "Life history","Net","Hook","Trawl","Trap","Vu"),
                 colwidths = c(1, 7,4,4,4,4,1))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = 'Table A.3.3 PSA scores for the species considered in the assessment (NB, catsharks, sawsharks, thresher sharks, wobbegongs, angel sharks, banjo rays, wedgefishes, stingrays, and ‘rays and skates’ were aggregated and assessed as at the group level due to unreliable species reporting for some fisheries). MA, maximum age; AM, age at maturity; FE, fecundity; MS, maximum size; SM, size at maturity; RS, reproductive strategy; TL, trophic level; Av, availability; En, encounterability; Sel, selectivity; PCM, post-capture mortality; VU, vulnerability.',
              autonum = NULL)%>%
  flextable::align(align = "center", part = "all")%>%
  flextable::align(j=1,align='left',part='all')  

```
<!---BLOCK_LANDSCAPE_STOP--->
## Environment
The spatio-temporal distribution of CPUE for the TDGDLF was combined with annual sea surface temperature anomalies to determine possible environment-driven changes in species distributions. In addition, the vulnerability of sharks and rays to climate change was extracted from published semi-quantitative risk assessments [@chinIntegratedRiskAssessment2010; @Walker2021].

```{r, results = 'asis'}
Paz=handl_OneDrive("Analyses/Catch and effort/Outputs/Spatial_Climate change/elasmos")
plots <- list.files(Paz) 
plots=plots[-grep(paste(c('Eagle','Pencil','Common sawshark','Shortfin mako'),collapse='|'),plots)]  #no stand. cpue
for(i in 1:length(plots))
{
  filename <- file.path(paste(Paz, plots[i],sep='/'))
  Titl=gsub("\\..*", "", plots[i])
  if(Titl=="Dusky whaler") Titl="Dusky shark"
  if(Titl=="Hammerhead sharks") Titl="Hammerheads"
  cat(sprintf('![%s](%s)', paste('Figure A.3.1.',i,'. ',Titl,'. Spatial distribution of nominal catch per unit of effort (CPUE, kg per km gillnet hour) for the TDGDLF. The size of the bubble is proportional to the CPUE value. Also shown are the annual sea surface temperature anomalies.',sep=''), 
               filename), sep = '\n')
  cat("\n")
}
```


<br>



# Appendix 4. Catch
A breakdown of the combined reported total catch of sharks and rays taken by Western Australian fisheries is shown in Figure A.4.1. Since 1975-76, >40 fisheries have historically reported catches of shark and ray species in statutory fishing returns but ~3/4 of the total shark and ray catch taken in WA originates from the state's shark fisheries, particularly gillnet fisheries.

```{r, echo=FALSE}
path=paste(Model.outputs.an,'Proportional catch_WA_combined.tiff',sep='')
```
![Figure A.4.1. Proportional shark and ray catch for Western Australia fisheries. The top panel shows the proportion by method for shark and non-shark fisheries. The bottom panel is a word cloud of the different fisheries interacting with sharks and rays. The southern and northern shark fisheries are shown in green and orange, respectively, whereas non-shark fisheries are shown in black. The size of the word is proportional to the square root of the total catch.](`r path`){width=800px}


For each assessed species, reported and reconstructed time series of catch by fishery is shown in Figure A.4.2 for indicator species and Figure A.4.3 for non-indicator species.

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Catch_all_species_Indicator.sp.tiff',sep='')
```
![Figure A.4.2. Catch time series for indicator species. Recreational, WA reconstructed recreational harvest (boat-based, shore-based and charter boats); TEP, WA Reported Threatened, Endangered and Protected species catches; WRL, WA reconstructed West Coast Rock lobster Fishery; NT, Northern Territory's commercial landings; SA MSF, South Australia’s Marine Scalefish Fishery landings; GAB Trawl, Great Australian Bight Trawl Fishery landings; WTB, Western Tuna and Billfish Fishery landings; Taiwan, reconstructed catches by Taiwanese vessels in WA waters; Indonesia, reconstructed catches by Indonesian vessels in WA waters.](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Catch_all_species_Other.sp.tiff',sep='')
```
![Figure A.4.3. Catch time series for non-indicator species. Recreational, WA reconstructed recreational harvest (boat-based, shore-based and charter boats); TEP, WA Reported Threatened, Endangered and Protected species catches; WRL, WA reconstructed West Coast Rock lobster Fishery; NSW Fisheries, New South Wales Fisheries commercial landings; SA MSF, South Australia’s Marine Scalefish Fishery landings; GAB Trawl, Great Australian Bight Trawl Fishery landings; WTB, Western Tuna and Billfish Fishery landings; Taiwan, reconstructed catches by Taiwanese vessels in WA waters; Indonesia, reconstructed catches by Indonesian vessels in WA waters.](`r path`){width=800px}

<br>

# Appendix 5. Effort and management dynamics for shark fisheries
For the TDGDLF and NSF, due to the overall rapid increase in catches between the mid 1970s and 1990s, and sustainability concerns around the general vulnerability of shark species, various management measures and data collection initiatives have been introduced since the late 1980s to reduce effort, and consequently catches (Figure A.5).

```{r, echo=FALSE}
path=paste(handl_OneDrive('Management/Sharks/Timeline management measures/'),'Effort.management_TDGDLF & NSF.tiff',sep='')
```
![Figure A.5. Total reported effort for the TDGDLF (top pane) and NSF (bottom pane). Also shown are the management measures and data collection initiatives implemented since the start of these fisheries. Effort is shown in a logarithmic scale. Red, data collection; Blue, spatial closures; Green, other management measures.](`r path`){width=800px}

<br>


# Appendix 6. Spatio-temporal changes in fishing effort and reported catch
Most of the shark and ray catch taken in WA originates from the state's shark fisheries (Figure A.4.1). Hence, analysis of spatio-temporal catch and effort dynamics focused on these fisheries. 

The spatio-temporal distribution of effort is shown in Figure A.6.1. Effort values are displayed in relative terms (i.e., relative to the maximum) because the NSF used predominantly demersal longline fishing gear and the TDGDLF used predominantly demersal gillnet fishing gear.

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Spatio.temporal.effort.tiff',sep='')
```
![Figure A.6.1. Reported effort dynamics for WA’s shark fisheries. Heatmap of the distribution of relative effort (hook days for the NSF and km gillnet days for the TDGDLF) by 5-year intervals and 60 nm blocks.](`r path`){width=800px}


The spatio-temporal distribution of reported catch for the TDGDLF is shown in Figures A.6.2 and A.6.3 for indicator and non-indicator species, respectively. 

For each assessed species, a single Consequence and Likelihood value was selected based on the spatial and temporal dynamics of effort, the total number of blocks fished and the proportion of blocks fished with catch within each species core area. Market dynamics and species protection were also taken in consideration. For angel sharks, most of the catch has been taken by the Great Australian Bight (GAB) Trawl Fishery (managed by the Commonwealth) whereas for copper shark a large proportion of the catch has been taken by South Australia's Marine Scale Fishery (SA MSF). No spatio-temporal catch distribution data were available from these fisheries. However, effort has substantially decreased in recent years due to the introduction of a range of management measures [@afmaHarvestStrategyFramework2009; @steerAssessmentSouthAustralian2020; @nitschkeSpatiotemporalVariabilityDemersal2022]. Similarly, fisheries operating under WA's jurisdiction have also gone through a range of management measures aimed at reducing fishing effort [@Gaughan2019].


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Spatio.temporal.catch_Indicator.sp.tiff',sep='')
```
![Figure A.6.2. Reported catch dynamics for the TDGDLF (indicator species). Proportion of blocks fished with reported catch (primary y axis, red dots) and total number of blocks fished (secondary y axis, grey line) within the species core area. The inset shows the species distribution within WA (green polygon) reported in https://www.cmar.csiro.au/data/caab and the species’ core area (black dots) within the fished zones. Core area was defined as the spatial blocks accounting for 95% of the catch.](`r path`){width=800px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Spatio.temporal.catch_Other.sp.tiff',sep='')
```
![Figure A.6.3. Reported catch dynamics for the TDGDLF (non-indicator species). Proportion of blocks fished with reported catch (primary y axis, red dots) and total number of blocks fished (secondary y axis, grey line) within the species core area. The inset shows the species distribution within WA (green polygon) reported in https://www.cmar.csiro.au/data/caab and the species’ core area (black dots) within the fished zones. Core area was defined as the spatial blocks accounting for 95% of the catch.](`r path`){width=800px}

<br>

# Appendix 7. Commercial catch length composition
Sex-specific length composition from the northern (NSF) and southern  (TDGDLF) shark fisheries are available for a subset of species collected by onboard observers. The sampling protocol is described in Braccini and Taylor [-@Braccini2016b]. No data are currently available on catch age-composition. 

```{r, results='asis', fig.width = 6, fig.height=6}
Species.with.length.comp<- read.csv(paste(Model.outputs.RAR,'/Species.with.length.comp.csv',sep=''))
Species.with.length.comp=Species.with.length.comp%>%pull(x)
sort.species.id=match(paste('1.',Indicator.species,sep=''),Species.with.length.comp)
only.indicator=FALSE
if(only.indicator)
{
  Species.with.length.comp=Species.with.length.comp[sort.species.id]
}
if(!only.indicator)
{
  Species.with.length.comp=c(Species.with.length.comp[sort.species.id],Species.with.length.comp[-sort.species.id])
}
Show.other.species=c("Copper shark","Angel sharks")  
for(i in 1:length(Species.with.length.comp))
{
  Titl=gsub(".*\\.", "", Species.with.length.comp[i])
  disfile=paste(Model.outputs.an,paste(Species.with.length.comp[i],this.yr,'1_Inputs/Visualise data',sep='/'),sep='')
  
  file.ls <- list.files(path=disfile, pattern=paste0("Size.comp",".*\\.tif"),all.files = TRUE,  
                        full.names = TRUE, recursive = TRUE)
  file.ls.names<- list.files(path=disfile, pattern=paste0("Size.comp",".*\\.tif"))
  file.ls=file.ls[-grep(paste(c("_dist","Catch curve"),collapse='|'),file.ls)]
  file.ls.names=file.ls.names[-grep(paste(c("_dist","Catch curve"),collapse='|'),file.ls.names)]
  file.ls.names=gsub("Size.comp.","",gsub('.tiff','',file.ls.names))
  if(!Titl %in%Show.other.species)
  {
    iaid=match("Other",file.ls.names)
    file.ls=file.ls[-iaid]
    file.ls.names=file.ls.names[-iaid]
  }
  for(l in 1:length(file.ls))
   {
     Nme=file.ls.names[l]
     if(Nme=='Other')
     {
      if(Titl=="Copper shark") Nme='SA MSF' 
      if(Titl=="Angel sharks") Nme='GAB Trawl' 
     }
     if(Nme=='Pilbara.trawl') Nme='Pilbara Fish Trawl'
     if(Nme=="TDGDLF") Nme= paste(Nme, 'by zone and mesh size (WC, West Coast; Zn1, Zone 1; Zn2, Zone 2)')
     if(!Nme=="Other")
     {
      cat(sprintf('![%s](%s)', paste('Figure A.7.',i,'.',l,'. ',Titl,'. Observed size composition in the catches of the ',Nme,'.',sep=''), 
                 file.ls[l]), sep = '\n')
     cat("\n")
     }
   }
  cat("\n")
}
```


<br>

# Appendix 8. Standardised catch per unit of effort
Abundance indices are available for a subset of species from standardising the commercial catch and effort from the TDGDLF (kg per km gillnet-hours) and from the longline abundance survey conducted in north-western WA (number of individuals per 50 hook-hours). The standardisation process is described in Braccini et al. [-@bracciniSpatialTemporalPatterns2021] for the fishery-dependent abundance index and in Braccini et al. [-@Braccini2020] for the fishery-independent abundance index. Brown dots indicate the years not considered when conditioning the stock assessments described in Appendices 11 and 12. 

```{r, results='asis'}
Species.with.abundance<- read.csv(paste(Model.outputs.RAR,'/Species.with.abundance.csv',sep=''))
Species.with.abundance=Species.with.abundance%>%pull(x)
if(!"1.Dusky shark"%in%Species.with.abundance) Species.with.abundance=c(Species.with.abundance,"1.Dusky shark")
if(!"1.Spinner shark"%in%Species.with.abundance) Species.with.abundance=c(Species.with.abundance,"1.Spinner shark")
Species.with.abundance=sort(Species.with.abundance)
sort.species.id=match(paste('1.',Indicator.species,sep=''),Species.with.abundance)
Species.with.abundance=c(Species.with.abundance[sort.species.id],Species.with.abundance[-sort.species.id])

for(i in 1:length(Species.with.abundance)){
  Titl=gsub(".*\\.", "", Species.with.abundance[i])  
       cat(sprintf('![%s](%s)', paste('Figure A.8.',i,'. ',Titl,'. Standardised catch per unit of effort (CPUE, mean and 95% CI).',sep=''), 
                 paste(Model.outputs.an,paste(Species.with.abundance[i],this.yr,
                      '1_Inputs/Visualise data/All cpues.tiff',sep='/'),sep='')), sep = '\n')
  cat("\n")
}

```

<br>

# Appendix 9. Demographic modelling
Demographic modelling was used to construct priors for the maximum intrinsic rate of population increase, *r* [@McAllister2001], a required input for the population models described below. Life history information  and Leslie Matrices were combined in a Monte Carlo procedure (10,000 iterations) to derive *r* priors as per described in Braccini et al. [-@bracciniDisplayingUncertaintyBiological2015]. A birth-pulse (i.e. a discrete breeding season), post-breeding census and a yearly time step were assumed (e.g., Aires-da-Silva and Gallucci [-@Aires-da-Silva2007]). The *popbio* R package [@stubbenEstimatingAnalyzingDemographic2007] was used to solve the Leslie Matrix.
Information on female litter size, maximum age, age at 50% maturity (A~50~), reproductive cycle length, growth, and the assumed distributions for these life history parameters were sourced from the literature (Table A.9.1). Instantaneous natural mortality at age (M~*a*~) was derived using the unified M estimation method developed by Dureuil et al. [-@dureuilUnifiedNaturalMortality2021] for elasmobranchs. Two scenarios were considered, one using a constant M value, set at the adult M, and one using the M~*a*~ vector.
<!---BLOCK_LANDSCAPE_START--->
<!-- Life history table -->
```{r,echo=FALSE,ft.align="left"}
Tab<- read.csv(paste(Model.outputs.RAR,'/Table 1. Life history pars.csv',sep=''))
Tab=Tab%>%
  mutate(a_w8t=formatC(a_w8t, format = "e", digits = 3))
names(Tab)=gsub("\\.", "", names(Tab), perl=TRUE)
disnames=c("Kyear1","FL_infcm","MaxAgeyear","Age50Matyear","TL50matcm","TL95matcm","Littersize",  
           "Repcycleyear","a_w8t","b_w8t")
names(Tab)[match(disnames,names(Tab))]=c("k","FLinf","Max age","Age 50% Mat","TL 50 Mat",
                                         "TL95 Mat","Litter size",  
                                         "Rep. cycle","a","b")
flextable(Tab)%>%
  width(j='Species',width=3,unit='cm')%>%
  width(j='a',width=1.8,unit='cm')%>%
  width(j=5,width=3,unit='cm')%>%
  width(j=8,width=3.5,unit='cm')%>%
  width(j=9,width=2.5,unit='cm')%>%
  align(j=1,align='left',part='all')%>%
  align(j=2:11,align='center',part='all')%>%
  fontsize(size=9, part = "all")%>%
  #compose(i=1,j=2,part='header',value=as_paragraph('K (year',as_sup('-1'),')'))%>%
  flextable::compose(i=1,j=3,part='header',value=as_paragraph('FL',as_sub(stri_unescape_unicode('\\u221E'))))%>%
  flextable::compose(i=1,j=5,part='header',value=as_paragraph('Age',as_sub('50')))%>%
  flextable::compose(i=1,j=6,part='header',value=as_paragraph('TL',as_sub('50')))%>%
  flextable::compose(i=1,j=7,part='header',value=as_paragraph('TL',as_sub('95')))%>%
  set_caption(caption = 'Table A.9.1. Life history information used for the calculation of r and h. k, growth coefficient; FL, fork length (cm); TL, total length (cm); Age50; age at 50% maturity (year); TL50 and TL95, total length at 50% and 95% maturity (cm); Rep. cycle, reproductive cycle length (year); a and b, parameters of the length-weight relationship (in cm and kg, respectively).',
              autonum = NULL)
```
<!---BLOCK_LANDSCAPE_STOP--->

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Prior_r_Indicator.sp.tiff',sep='')
```
![Figure A.9.1. Distribution of maximum intrinsic rate of population increase (r) derived from the demographic analysis for indicator species.](`r path`){width=800px}

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Prior_r_Other.sp.tiff',sep='')
```
![Figure A.9.2. Distribution of maximum intrinsic rate of population increase (r) derived from the demographic analysis for non-indicator species.](`r path`){width=800px}

# Appendix 10. Catch-only modelling
Catch-only models (COMs) were used to determine catch sustainability based on the comparison of catch trajectories and model-inferred MSY. Three COMs were used to account for model structural uncertainty and integrate across alternative processes: DBSRA, depletion-based stock reduction analysis  [@Dick2011], Catch-MSY [@Martell2012], and SSS, Simple Stock Synthesis [@copeImplementingStatisticalCatchatage2013]. The selected methods are within the recommended stock assessment toolbox for data-limited assessments, make use of available life history information and full catch trajectories, allow specifying initial and final depletion conditions, and are implemented in R. 
To inform model dynamics, we used species-specific life history (Table A.9.1) and stock-specific depletion information derived from the known history of management and fishery dynamics in WA (Figures A.5 and A.6) because catch declines are not only due to population declines driven by overfishing but also due to environmental conditions affecting productivity and/or fisheries regulation explicitly designed to reduced catches [@Ovando2022]. 

## DB-SRA
DB-SRA, available in the *fishmethods* R package [@Nelson2021], simulates a stock’s biomass trajectory using a catch time series and a delay-difference model parametrised as function of M, A~50~, F~MSY~/M (the ratio of F~MSY~, the fishing mortality associated with MSY, and M), B~MSY~/K (the ratio of B~MSY~, the biomass level associated with the MSY, and K, stock carrying capacity), and the expected depletion at the start (B~0~/K) and end (B~f~/K) of the catch time series. Monte Carlo simulations (10,000 iterations; preliminary analysis showed similar results from a larger number of iterations), were used to draw random parameter values from specified prior distributions. Using the catch history, a biomass trajectory is calculated based on the sampled input parameters. The trajectory and corresponding input parameters are retained if the biomass trajectory does not go below zero. The parameter values and distributions used in this study are presented in Table A.10.1. Distributions for M and A~50~ were obtained from the demographic modelling outlined above assuming a lognormal distribution (e.g., Sweka et al. [-@swekaApplicationDepletionBasedStock2018]) with CV calculated as the standard deviation (SD) over the mean. F~MSY~/M had a lognormal distribution with a mean of 0.41 [@zhouLinkingFishingMortality2012] and a CV of 0.1. B~MSY~/K had a beta distribution with mean calculated based on Cortes [-@Cortes2012] and CV of 0.1. B~0~/K had uniform distributions ranging [0.95, 1]. A high B~0~/K level was assumed because the start of the catch time series is considered to be the start of any type of commercial/recreational shark fishing in WA so stocks are assumed to be at around carrying capacity. For B~f~/K, it is unlikely that stocks in WA are at very low depletion levels given that total catches have been in the order of tens of tonnes or less per year for most species and for some species catches have been negligible over multiple generations. Hence, for most species B~f~/K had uniform distributions ranging [0.2,0.95] so as not to restrict the model. For some species with high productivity and negligible exploitation in recent years the lower bound was increased to allow model convergence. Finally, K had a uniform distribution ranging between five and 50 times the species’ maximum annual catch (e.g., Andrade [-@Federal2017]). Model parameter inputs are shown in Table A.10.1. 


## Catch-MSY
Catch-MSY simulates a stock’s biomass trajectory using a catch time series and a surplus production model parametrised as function of *r*, K, B~0~/K, and B~f~/K. As for DB-SRA, catch time series and input parameter values, drawn from a Monte Carlo procedure, are used to reconstruct biomass trajectories, retaining the combinations of *r* and K values that did not cause the stock to go extinct or above K [@Martell2012]. Following Dichmont et al. [-@Dichmont2021], we used Haddon’s parametrisation of Catch-MSY available from the *datalowSA* R package [@Haddon2019]. Rather than sampling *r* values from the prespecified default productivity ranges, a uniform distribution was used with lower and upper ranges set at the 10% and 90% percentiles, respectively, of the *r* distributions derived from demographic modelling. Catch-MSY allows incorporating process error to simulate stochasticity in biomass dynamics. This is considered to be much lower for sharks than for broadcast spawners such as teleosts and marine invertebrates. Hence, process error was set at 1e^-3^ as catch-only methods yield considerably uncertain estimates by nature. The K, B~0~/K, and B~f~/K priors were modelled as for the DB-SRA models. Model parameter inputs are shown in Table A.10.2.


## Simple Stock Synthesis (SSS)
SSS is based on the Stock Synthesis stock assessment model [@methotStockSynthesisBiological2013], providing an age-structured alternative to the other COMs. SSS uses a Monte Carlo approach to draw samples of key input parameters, which are prespecified in the Stock Synthesis model, and estimates only the log of unexploited equilibrium recruitment (lnRo). Input parameters drawn from prior distributions included M, B~0~/K, B~f~/K, and steepness, *h*, the fraction of recruitment from a virgin population obtained when the population is at 20% virgin level. Other input parameters were prespecified at the midpoint of the reported range used for demographic modelling (Table A.9.1). Values of M were drawn from a lognormal distribution with mean and SD set at the mean and SD of the empirical M distribution obtained from the demographic analysis. Values of B~0~/K and B~f~/K were drawn from the prior distributions used for DB-SRA. Steepness (mean and SD) was calculated using the analytical solution derived by Brooks et al. [-@Brooks2010] and the life history information used in the demographic analysis. For great (*Sphyrna mokarran*) and smooth (*S. zygaena*) hammerheads and tiger (*Galeocerdo cuvier*), zebra (*Stegostoma tigrinum*) and whiskery sharks, the derived *h* values were 0.77, 0.66, 0.70, 0.71 and 0.60, respectively. These high *h* values resulted from the combination of relatively high fecundity and substantial longevity (greatest for hammerheads and tiger shark), low juvenile mortality and early age at maturity. For sawsharks, green sawfish (*Pristis zijsron*) and milk shark (*Rhizoprionodon acutus*), the derived *h* values were 0.31, 0.36 and 0.44, respectively, which are considered likely to represent underestimates for these relatively productive species. Hence, alternative *h* values were also considered for sensitivity analysis. These alternative *h* values were derived from a linear model fitted to the *r* and *h* values calculated for the other assessed species (Figure A.10.1). For SSS, random values of *h* were drawn from a truncated beta distribution with mean and SD derived from the calculation of *h* described above and boundaries set at 0.3 and 0.8, respectively. The lower bound was set at 0.3 on the basis of the evolutionary principle that a minimum recruitment compensation is needed for population persistence given recruitment variability [@He2006]. The upper bound was set at the average *h* value for blue shark (*Prionace glauca*), considered to be one of the most productive shark species [@Cortes2016a; @Kai2018]. Model runs were retained if the difference between observed and predicted B~f~/K values was ≤0.01 [@copeImplementingStatisticalCatchatage2013]. Model parameter inputs are shown in Table A.10.3.

<!---BLOCK_LANDSCAPE_START--->
<!-- DBSRA scenarios -->
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_DBSRA_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_DBSRA_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
      arrange(Species)%>%
  dplyr::select(-Scenario)%>%
  mutate(bmsyk.mean=round(bmsyk.mean,2))
flextable(Tab)%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(j='Species',width=2.5,unit='cm')%>%
  width(j=c(3,8),width=2,unit='cm')%>%
  width(j=c(11),width=1,unit='cm')%>%
  width(j=c(14,17),width=1.5,unit='cm')%>%
  width(j=2,width=2,unit='cm')%>%
  width(j=c(2,5:6,10,13,15:16,18:19),width=1,unit='cm')%>%
  width(j=6:7,width=1.35,unit='cm')%>%
  width(j=c(4,9,12),width=1.2,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=1,align='left',part='all')%>%
  set_header_labels(
      Species="",AgeMat='',M.dist='Dist.',Mmean='Mean',M.CV='CV',
      Klow="Low",Kup="Up",fmsym.dist="Dist.",fmsy.m="Mean",fmsym.CV="CV",
      bmsyk.dist="Dist.",bmsyk.mean="Mean",bmsyk.CV="CV",
      b1k.dist="Dist.",b1k.low="Low",b1k.up="Up",
      btk.dist="Dist.",btk.low="Low",btk.up="Up")%>%
  add_header_row(values = c("Species","Age at","M","K","FMSY","BMSY/K","Bo/K","Bf/K"),
                 colwidths = c(1,1,3,2,3,3,3,3))%>%
  flextable::compose(i=1,j=2,part='header',value=as_paragraph('Age',as_sub('50')))%>%
  flextable::compose(i=1,j=8,part='header',value=as_paragraph('F',as_sub('MSY')))%>%
  flextable::compose(i=1,j=11,part='header',value=as_paragraph('B',as_sub('MSY'),'/K'))%>%
  #compose(i=1,j=14,part='header',value=as_paragraph('B',as_sub('o'),'/K'))%>%
  #compose(i=1,j=17,part='header',value=as_paragraph('B',as_sub('f'),'/K'))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.10.1. Input parameter values for the DBSRA models. Age50, age at 50% maturity (year); M, natural mortality; Dist., distribution; K, carrying capacity (tonnes); Low, lower bound; Up, upper bound; FMSY, Fishing mortality at MSY; BMSY, Biomass at MSY; Bo, initial biomass; Bf, final biomass.",
              autonum = NULL)
```

\newpage
<!-- CMSY scenarios -->
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_CMSY_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_CMSY_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
      arrange(Species)%>%
  dplyr::select(-Scenario)
flextable(Tab)%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=2,unit='cm')%>%
  width(j='Species',width=2.5,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=1,align='left',part='all')%>%
  set_header_labels(
      Species="",r.low='Low',r.up='Up',Klow='Low',Kup='Up',
      Bo.low='Low',Bo.hi='Up',Bf.low='Low',Bf.hi='Up',Proc.error='')%>%
  add_header_row(values = c("Species","r","K","Bo/K","Bf/K","Proc. error"),
                 colwidths = c(1,2,2,2,2,1))%>%
  flextable::compose(i=1,j=2,part='header',value=as_paragraph(as_i('r')))%>%
  #compose(i=1,j=6,part='header',value=as_paragraph('B',as_sub('o'),'/K'))%>%
  #compose(i=1,j=8,part='header',value=as_paragraph('B',as_sub('f'),'/K'))%>%
  #compose(i=1,j=10,part='header',value=as_paragraph(stri_unescape_unicode('\u03c3'),as_sup('2')))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.10.2. Input parameter values for the CMSY models. r, maximum intrinsic rate of population increase; K, carrying capacity (tonnes); Low, lower bound; Up, upper bound; Bo, initial biomass; Bf, final biomass; Proc. error, process error.",
              autonum = NULL)
```

\newpage
<!-- SSS scenarios -->
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_SSS_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 2. Catch only_scenarios_SSS_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
      arrange(Species)%>%
  dplyr::select(-c(F_ballpark))%>%
  mutate(Mmean=round(Mmean,2),
         h.mean=round(h.mean,2),
         Msd=round( Msd,3))%>%
  relocate(Msd,.after=Mmean)
flextable(Tab)%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=1.25,unit='cm')%>%
  width(j='Species',width=2.5,unit='cm')%>%
  width(j='Scenario',width=2,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=1,align='left',part='all')%>%
  set_header_labels(
      Species="",Scenario="",Mmean="Mean",Msd="SD",h.mean="Mean",h.sd="SD",Initial.bio.low="Low",Initial.bio.up="Up",
      Final.bio.low="Low",Final.bio.up="Up",Ln_R0_min="Low",Ln_R0_max="Up")%>%
  add_header_row(values = c("Species","Scenario","M","h","Bo/K","Bf/K","ln(Ro)"),
                 colwidths = c(1,1,2,2,2,2,2))%>%
 # compose(i=1,j=6,part='header',value=as_paragraph('B',as_sub('o'),'/K'))%>%
 #  compose(i=1,j=8,part='header',value=as_paragraph('B',as_sub('f'),'/K'))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.10.3. Input parameter values for the SSS models. M, natural mortality; h, steepness; Low, lower bound; Up, upper bound; Bo, initial biomass; Bf, final biomass; lnRo, log of unexploited equilibrium recruitment.",
              autonum = NULL)
```

<!---BLOCK_LANDSCAPE_STOP--->

The COMs considered here represent alternative states of nature; hence, an ensemble was constructed as model ensembles can streamline stock assessments and are typically more accurate and less biased than individual COM estimates [@jardimOperationalizingEnsembleModels2021]. For each COM, random samples of accepted trajectories and inferred MSY values were drawn and the 60% and 95% percentiles were calculated. When drawing random samples, SSS was given double the weight than the other COMs given the higher biological realism (e.g. sex- and age-structure, stock-recruitment relationship) of SSS.

<br>

```{r, echo=FALSE}
path=paste(Model.outputs.an,'Steepness_vs_r.tiff',sep='') 
```
![Figure A.10.1. Linear regression of maximum intrinsic rate of population increase (r) and steepness (h).](`r path`){width=800px}

<br>

```{r,echo=FALSE,ft.align="left"}
#quantities of interest
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 4. Catch.only_catch_vs_MSY_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 4. Catch.only_catch_vs_MSY_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
  arrange(Species)%>%
  dplyr::select(Species,MSY_Median,MSY_Lower.95,MSY_Upper.95)%>%
  mutate_at(vars('MSY_Median','MSY_Lower.95','MSY_Upper.95'), round, 2)%>%
  rename(Median=MSY_Median,
         Lower.95=MSY_Lower.95,
         Upper.95=MSY_Upper.95)%>%
  relocate(Species,Lower.95)

Tab%>%
  flextable()%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=2.5,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=c('Species'),align='left',part='all')%>%
  width(j='Species',width=3.5,unit='cm')%>%
  set_caption(caption = "Table A.10.4. Inferred MSY values from the COM ensemble for the base case scenario.",
              autonum = NULL)


```

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Catch_vs_MSY_catch.only_Indicator.sp.tiff',sep='')
```
![Figure A.10.2. Time series of total catch (blue line) and MSY (black line) with 60% and 95% confidence intervals (dark grey and light grey shaded areas, respectively) inferred from the catch-only model ensemble for indicator species. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Catch_vs_MSY_catch.only_Other.sp.tiff',sep='')
```
![Figure A.10.3. Time series of total catch (blue line) and MSY (black line) with 60% and 95% confidence intervals (dark grey and light grey shaded areas, respectively) inferred from the catch-only model ensemble for non-indicator species. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

<br>

# Appendix 11. State-space surplus production modelling
State-space surplus production models were fitted to catch and abundance series derived from standardised commercial catch and effort [@bracciniSpatialTemporalPatterns2021] and/or a fishery-independent survey [@Braccini2020] using JABBA, Just Another Bayesian Biomass Assessment [@winkerJABBAJustAnother2018]. The survey collects abundance data as shark numbers so it was converted to biomass by multiplying numbers by the reported average annual weight [@Braccini2020]. A full description of the JABBA model (state-space implementation, prior specification options, diagnostic tools, etc.) is available in Winker et al. [-@winkerJABBAJustAnother2018]. JABBA is implemented in R with a JAGS interface [@Plummer2003] to estimate the Bayesian posterior distributions using Markov Chains Monte Carlo (MCMC) simulations. A Pella-Tomlinson production function was used by setting the inflection point B~MSY~/K and converting this ratio into a shape parameter *m* that determines at which B/K ratio maximum surplus production is attained. Schaefer and Fox production functions were considered as sensitivity tests. We used two MCMC chains each of 30,000 iterations, sampled with a burn-in period of 5,000 iterations and thinning rate of five iterations (e.g., Winker et al. [-@Winker2019]). Model diagnostics included Bayesian fit diagnostics (visualization of the MCMC chains, posterior densities, autocorrelation and the Gelman et al. [-@Gelman2003] and Geweke [-@Geweke1992] diagnostics as implemented in the *ggmcmc* R package [@FernandezMarin2016]) and model goodness-of-fit, retrospective analysis and prediction skill as implemented in the *ss3diags* R package [@carvalhoCookbookUsingModel2021]. Key priors required are *r*, K and B~0~/K whereas priors for B~MSY~/K, and process error were also specified (Table A.11.1). A lognormal distribution was used for *r* and K and a beta distribution was used for B~0~/K as it is considered to have better convergence properties than uniform priors [@winkerJABBAJustAnother2018]. For *r*, the mean and CV were obtained from the demographic modelling. For K, the mean was set at the average of the uniform distribution used in the COMs, and the CV was prespecified at 200% (Winker et al., 2019). A uniform distribution, prespecified at the range used for the COMs, was considered as a sensitivity. For B~0~/K, the mean was prespecified to the average of the lower and upper bounds of the B~0~/K distribution used in the COMs and the CV was calculated as the SD of those distribution over the mean. Process error was prespecified at 0.05, which assumes low stochastic variation in biomass given sharks productivity (Winker et al., 2019). We used the same B~MSY~/K values used in the DB-SRA modelling. 

Standardisation of CPUE can yield very small CVs (as observed for the analysed species); hence, the minimum CV used in the models was extracted from fitting a simple smoother to the CPUE data (on the natural log scale) and estimating the residual variance [@Francis2011]. A thorough description of this method can be found in Courtney et al. [-@courtneyStockSynthesisSS32017]. 

The construction of JABBA models was possible for a subset of species (copper, dusky, gummy, milk, sandbar, tiger and whiskery sharks) due to availability of reliable survey and/or commercial CPUE series. 

For spinner shark, the TDGDLF CPUE was considered not likely be representative of abundance as it tracked catch trends (Figures A.4.3, A.8.7). 

For dusky shark, the reliability of the TDGDLF CPUE as an index of abundance is unclear. Due to the tight gillnet selectivity schedule of the mesh sizes used in the TDGDLF, the catch comprises mostly of 0+ and 1 year old individuals, potentially representing a recruitment index. However, between 1975 and 1985 catches of the NSF and ‘other’ fisheries (which caught adult individuals) moderately increased from 36 to 74 tonnes but the TDGDLF monthly CPUE greatly decreased (75% reduction). Also, overall fishery catches substantially increased after 1985 (Figure A.4.2) and these higher catch levels were able to be sustained, suggesting that recruitment was unlikely to have markedly declined between 1975 and 1985 as suggested by the TDGDLF CPUE (Figure A.8.2). Hence, if the TDGDLF CPUE was a recruitment index, a substantial recruitment reduction between 1975 and 1985 within the context of the species exploitation history would be biologically unfeasible. Furthermore, since the late 1980s, the TDGDLF CPUE fluctuated showing no overall trend (hence no trend in recruitment) despite catches (of adults and juveniles) substantially increasing to a history peak in the late 1980s and then substantially decreasing due to management-driven effort reductions (Figures A.4.2, A.5). The fishery-independent survey is highly uncertain as only a few individuals are caught each year due to nature of the survey (20 fishing stations revisited annually) which is designed to target mostly sandbar sharks (Figure A.8.2). Hence, abundance series were considered unreliable and JABBA models were not built for dusky shark.

For sandbar shark, the TDGDLF monthly CPUE series was excluded from the analysis because it tracked the catch series (Figures A.4.2, A.8.4). Preliminary analysis using the JABBA model including this CPUE data showed similar population trajectories albeit considerably more uncertain, displaying a sudden peak in relative biomass in 1986, not consistent with sandbar shark productivity. Hence, the TDGDLF monthly CPUE was considered unreliable as an index of abundance and not used in the JABBA models.

For gummy shark, the years 1975 to 1985 of the monthly CPUE series were not included because the CPUE decreased by ~80% during this period following only a moderate increase in catches (Figures A.4.2, A.8.1) and the fleet was composed of a large number of vessels that fished mostly sporadically. 

For whiskery shark, changes in fishers' targeting behaviour (from whiskery to dusky sharks) in the early 1980s [@Simpfendorfer2000a] were modeled by the inclusion of two catchability coefficients for the periods 1975-1980 and 1984-2005. Finally, sensitivity tests were done to evaluate the effect of the assumed process error, *r*, production function and K prior (Table A.11.1).

Finally, for the TDGDLF CPUE derived from standardised commercial catch and effort data, the financial years 2007 and 2008 are considerably higher (~100%) than subsequent years for all species. This coincides with the transition of monthly to daily catch and effort reporting in WA which is considered likely to have distorted the calculation of effort variables [@bracciniSpatialTemporalPatterns2021] during this period of stable or decreasing catches. Hence, these two years were removed from the CPUE series as they are most likely reflecting changes in the reporting of the effort variables used in the standardisations.

<!---BLOCK_LANDSCAPE_START--->
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 7. JABBA CPUE_scenarios_Indicator.sp.csv',sep=''))
SP_ind=Tab1%>%pull(Species)
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 7. JABBA CPUE_scenarios_Other.sp.csv',sep=''))
SP_other=Tab2%>%pull(Species)

Tab=rbind(Tab1,Tab2)%>%
  arrange(Species)%>%
  mutate(r.mean=round(r.mean,2),
         PsiDist=capitalize(PsiDist),
         model.type=ifelse(grepl('Pella',model.type),'Pella-Tomlinson',model.type))%>%
  dplyr::select(-c(PsiDist,Bo.mean,Bo.CV))
if('Daily.cpues'%in%names(Tab))
{
  Tab=Tab%>%mutate(Daily.cpues=ifelse(is.na(Daily.cpues),'',Daily.cpues),
                   Daily.cpues=ifelse(Species%in%c('Dusky shark','Milk shark','Sandbar shark'),'N/A',Daily.cpues))
  flextable(Tab)%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=1.35,unit='cm')%>%
    width(j=c(3,6),width=1.75,unit='cm')%>%
    width(j=c(10,12),width=1.85,unit='cm')%>%
    width(j=c(13),width=3,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    width(j='Scenario',width=2,unit='cm')%>%
    align(align='center',part='all')%>%
    align(j=1,align='left',part='all')%>%
    set_header_labels(
      Species="",Scenario="",Rdist="Dist.",r.mean="Mean",r.cv="CV",
      Kdist="Dist.",K.mean="Mean",K.CV="CV",
      bmsyk.mean="",Proc.error="error",Ktch.CV="CV",Daily.cpues="CPUE",model.type="type")%>%
    add_header_row(values = c("Species","Scenario","r","K","Bmsyk","Process","Catch","Removed","Model"),
                   colwidths = c(1,1,3,3,1,1,1,1,1))%>%
    flextable::compose(i=1,j=3,part='header',value=as_paragraph(as_i('r')))%>%
    flextable::compose(i=1,j=9,part='header',value=as_paragraph('B',as_sub('MSY'),'/K'))%>%
    border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
    set_caption(caption = "Table A.11.1. Input parameter values for the Bayesian state-space surplus production models. r, maximum intrinsic rate of population increase; K, carrying capacity (tonnes); Dist., distribution.",
                autonum = NULL)
}else
{
  flextable(Tab)%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=1.35,unit='cm')%>%
    width(j=c(3,6),width=1.75,unit='cm')%>%
    width(j=c(10),width=1.85,unit='cm')%>%
    width(j=c(12),width=3,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    width(j='Scenario',width=2,unit='cm')%>%
    align(align='center',part='all')%>%
    align(j=1,align='left',part='all')%>%
    set_header_labels(
      Species="",Scenario="",Rdist="Dist.",r.mean="Mean",r.cv="CV",
      Kdist="Dist.",K.mean="Mean",K.CV="CV",
      bmsyk.mean="",Proc.error="error",Ktch.CV="CV",model.type="type")%>%
    add_header_row(values = c("Species","Scenario","r","K","Bmsyk","Process","Catch","Model"),
                   colwidths = c(1,1,3,3,1,1,1,1))%>%
    flextable::compose(i=1,j=3,part='header',value=as_paragraph(as_i('r')))%>%
    flextable::compose(i=1,j=9,part='header',value=as_paragraph('B',as_sub('MSY'),'/K'))%>%
    border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
    set_caption(caption = "Table A.11.1. Input parameter values for the Bayesian state-space surplus production models. r, maximum intrinsic rate of population increase; K, carrying capacity (tonnes); Dist., distribution.",
                autonum = NULL)
}
```
<!---BLOCK_LANDSCAPE_STOP--->

<!-- Display Sensitivity analysis--> 
 The resulting relative (total) biomass trajectories based on the scenarios considered are shown in Figures A.11.1 for indicator species and A.11.2 for non-indicator species. Overall, the different scenarios considered showed very similar trajectories and current depletion levels. On the basis of these comparisons, S1 was selected as the base case scenario.

<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_JABBA CPUE_Indicator.sp_sensitivity.tiff',sep='')
```
![Figure A.11.1. Sensitivity analysis showing the influence of the assumed value of r (S2), process error (S3) and K (S4), and type of production model (S5 & S6) for indicator species on the estimated relative (total) biomass (i.e. biomass over unfished biomass, solid line) with 95% credibility intervals (shaded area). Also shown are the reference points (red, limit; orange, threshold; green, target). The assumed r, process error and K values and type of production function for the different scenarios is shown in Table A.11.1.](`r path`){width=800px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_JABBA CPUE_Other.sp_sensitivity.tiff',sep='')
```
![Figure A.11.2. Sensitivity analysis showing the influence of the assumed value of r (S2), process error (S3) and K (S4), and type of production model (S5 & S6) for non-indicator species on the estimated relative (total) biomass (i.e. biomass over unfished biomass, solid line) with 95% credibility intervals (shaded area). Also shown are the reference points (red, limit; orange, threshold; green, target). The assumed r, process error and K values and type of production function for the different scenarios is shown in Table A.11.1.](`r path`){width=800px}

<br>


<!-- Display Bayesian convergence for S1--> 
For the base case scenario, Bayesian fit diagnostics showed good convergence properties for all species (Figure A.11.3). The posterior distributions for the two chains were aligned indicating convergence to the similar space. The trace plots showed good chain mixing. The running means diagnostic showed that chains quickly approached the posterior distribution and there was low autocorrelation beyond the first lag. Finally, the Gelman diagnostic values were close to the expected value of 1 [@Gelman2003] and the Geweke diagnostic values were within the expected range [-2,2] [@Geweke1992].
<br>

```{r, results='asis'}
# SP_ind=read.csv(paste(Model.outputs.RAR,"Table 7. JABBA CPUE_scenarios_Indicator.sp.csv",sep='/'))%>%
#           pull(Species)
# SP_other=read.csv(paste(Model.outputs.RAR,"Table 7. JABBA CPUE_scenarios_Other.sp.csv",sep='/'))%>%
#           pull(Species)
Species=unique(sort(c(SP_ind,SP_other)))
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.3.',i,'. ',Titl,'. Bayesian fit diagnostics for the base case scenario and the two MCMC chains. K, carrying capacity (tonnes); m, shape parameter of the production function; r, maximum intrinsic rate of population increase.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Fit_Bayesian convergence diagnostics.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display priors and posteriors for S1--> 
For the base case scenario, marginal posterior and prior distributions showed considerable overlap (Figure A.11.4), providing no evidence of severe prior misspecification and indicating that posteriors were largely informed by the priors. However, the K parameter was updated by the data for all species whereas the *m* and *r* parameters were updated by the data for tiger and whiskery sharks and to some extent sandbar shark.
<br>

```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.4.',i,'. ',Titl,'. Prior and posterior distributions for  the base case scenario. K, carrying capacity (tonnes); m, shape parameter of the production function; r, maximum intrinsic rate of population increase.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Prior.and.posterior.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display Fit to cpue series for S1--> 
For the base case scenario, the models fit the abundance series reasonably well with most fits falling within the 95% CI (Figure A.11.5). The runs tests provided no evidence to reject the hypothesis of randomly distributed residuals (Figure A.11.6). The goodness-of-fit was adequate for most species (RMSE≤30%, [@carvalhoCookbookUsingModel2021]) with the exception of milk and tiger sharks and to some extent copper shark, which showed larger RMSE values and residual boxplots, and an apparent trend (Figure A.11.7).
<br>

```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.5.',i,'. ',Titl,'. Fits to the abundance series for the base case scenario. Observed (white circles and 95% CIs) and predicted (solid line) abundance. The dark shaded grey area shows 95% credibility intervals and the light shaded grey area shows the 95% posterior predictive distribution intervals.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Fit_cpues.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display Runs test for S1--> 
```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.6.',i,'. ',Titl,'. Runs test for the base case scenario and fishery. Green panel, no evidence of lack of randomness of abundance series residuals (p≥0.05); Red panel, evidence of lack of randomness of abundance series residuals. The shaded area spans three standard errors and red circles denote years with greater residuals.',sep=''),
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,
                                           'JABBA CPUE/S1/Fit_runstest.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display dresiduals for S1--> 
```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.7.',i,'. ',Titl,'. Residual diagnostic plots for the base case scenario. Boxplot, median and quantiles of the residuals; solid black line, loess smoother through the residuals.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Fit_dresiduals.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display Retrospective analysis for S1--> 
For the base case scenario, there was an overall negligible retrospective pattern, producing consistent stock status estimates (Figure A.11.8) and therefore supporting the use of the base case model for making future projections. For all moderately-lived and long-lived species, the Mohn's-rho (ρ) statistic fell within the acceptable interval for long-lived species of -0.15 and 0.2 whereas for milk shark, a short-lived species, the ρ statistic fell within the acceptable interval for short-lived species of -0.22 and 0.3 [@carvalhoCookbookUsingModel2021].
<br>

```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.8.',i,'. ',Titl,'. Retrospective analysis results for the base case scenario done by removing one year at a time sequentially (n=5) and predicting relevant management quantities. Grey shaded area, 95% CI from the reference model.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Fit_retrospective.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<!-- Display Hindcasting for S1--> 
For the base case scenario, the hindcasting cross-validation results showed predictions fell within the 95% CI suggesting good predicting skills overall (Figure A.11.9). For copper and gummy sharks, the mean absolute scale error (MASE) score was > 1 (the reference level), which implies that the average model forecast is no better than a random walk [@carvalhoCookbookUsingModel2021]. For the other species, the MASE score was ≤ 1 (i.e., predictions are better than a random walk).
<br>

```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.11.9.',i,'. ',Titl,'. Hindcasting cross-validation results for the base case scenario showing one-year ahead forecast of abundance values, performed with 5 hindcast model runs relative to the expected value. Large color-coded circles, observations; grey shaded area, predicted 95% CI for the reference year; small color-coded circles, model predictions.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'JABBA CPUE/S1/Fit_hindcasting.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display parameter estimates for S1--> 
For the base case scenario, model parameter estimates and management quantities of interest are shown in Table A.11.2.
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 8. JABBA CPUE_estimates_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 8. JABBA CPUE_estimates_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
  filter(Scenario=='S1' & Parameter%in%c("Hmsy","K","MSY","r","SBmsy","m","P2021"))%>%
  arrange(Species)%>%
  dplyr::select(-c(Model, Scenario))%>%
  mutate_if(is.numeric, round, 2)%>%
  filter(!Parameter%in%c('Hmsy','SBmsy'))
  Tab%>%
    mutate(Parameter=ifelse(Parameter=='Hmsy','FMSY',
                     ifelse(Parameter=='SBmsy','BMSY',
                     ifelse(grepl("P",Parameter),'Current depletion',
                            Parameter))))%>% 
    flextable()%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=2.5,unit='cm')%>%
    width(j='Parameter',width=3.5,unit='cm')%>%
    set_header_labels(Parameter = "Parameter/Quantity")%>%
    set_caption(caption = "Table A.11.2. Summary of posterior parameter estimates and management quantities of interest with 95% Bayesian Credibility Intervals for the base case scenario. r, maximum intrinsic rate of population increase; K, carrying capacity (tonnes).",
                autonum = NULL)


```

<br>


<!-- Display biomass trajectories for S1--> 
For the base case scenario, the relative (total) biomass trajectories for indicator species are shown in Figure A.11.10.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_JABBA CPUE_Indicator.sp.tiff',sep='')
```
![Figure A.11.10. Estimated time series of relative (total) biomass (i.e. biomass over unfished biomass) for indicator species (60% and 95% credibility intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target) and the total catch series (broken blue line).](`r path`){width=800px}

For the base case scenario, the relative (total) biomass trajectories for non-indicator species are shown in Figure A.11.11.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_JABBA CPUE_Other.sp.tiff',sep='')
```
![Figure A.11.11. Estimated time series of relative (total) biomass (i.e. biomass over unfished biomass) for non-indicator species (60% and 95% credibility intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target) and the total catch series (broken blue line).](`r path`){width=800px}

<br>
<!-- Display F trajectories for S1--> 
For the base case scenario, the F trajectories for indicator species are shown in Figure A.11.12.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.series_JABBA CPUE_Indicator.sp.tiff',sep='')
```
![Figure A.11.12. Estimated time series of fishing mortality for indicator species (60% and 95% credibility intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

For the base case scenario, the F trajectories for non-indicator species are shown in Figure A.11.13.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.series_JABBA CPUE_Other.sp.tiff',sep='')
```
![Figure A.11.13. Estimated time series of fishing mortality for non-indicator species (60% and 95% credibility intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

<br>

<!-- Display Kobe plots for S1--> 
For the base case scenario, Kobe plots for indicator species are shown in Figure A.11.14.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_JABBA CPUE_WA_Fisheries_Indicator.sp.tiff',sep='')
```
![Figure A.11.14. Kobe plots showing the estimated trajectories of relative (total) biomass (i.e. biomass over unfished biomass, B0) and fishing mortality for indicator species. The red, orange, yellow and green panels indicate the areas below (above for fishing mortality) limit, between limit and threshold, between threshold and target, and above (below for fishing mortality) reference points.](`r path`){width=800px}

For the base case scenario, Kobe plots for non-indicator species are shown in Figure A.11.15.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_JABBA CPUE_WA_Fisheries_Other.sp.tiff',sep='')
```
![Figure A.11.15. Kobe plots showing the estimated trajectories of relative (total) biomass (i.e. biomass over unfished biomass, B0) and fishing mortality for non-indicator species. The red, orange, yellow and green panels indicate the areas below (above for fishing mortality) limit, between limit and threshold, between threshold and target, and above (below for fishing mortality) reference points.](`r path`){width=800px}

<br>


<!-- Display likelihoods for each Consequence--> 
For the base case scenario and the biomass reference points, the Probability for each Consequence level used in the Weight of Evidence assessment is shown in Table A.11.3.
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 9. JABBA CPUE_Current.depletion_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 9. JABBA CPUE_Current.depletion_Other.sp.csv',sep=''))
Tab=full_join(Tab1,Tab2,by=c("Range","Scenario","finyear","Model"))%>%
  filter(Scenario=='S1')%>%
  dplyr::select(-Scenario)%>%
  rename(Consequence=Range)%>%
  mutate(Consequence=case_when(Consequence=='<lim'~'Major',
                               Consequence=='lim.thr'~'High',
                               Consequence=='thr.tar'~'Moderate',
                               Consequence=='>tar'~'Minor'),
         Consequence=factor(Consequence,levels=c('Minor','Moderate','High','Major')))%>%
  gather(Species,Probability,-c(Consequence,finyear,Model))%>%
  spread(Consequence,Probability)%>%
  mutate(Species=gsub('\\.', ' ', Species))%>%
  rename(Year=finyear)%>%
  mutate(Year=as.character(Year))%>%
  arrange(Model,Species,Year)%>%
  dplyr::select(-Model)
  
 Tab%>% 
    flextable()%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=2,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    set_caption(caption = "Table A.11.3. Biomass reference points. Probability for each consequence levels used in the Weight of Evidence assessment.",
                autonum = NULL)%>%
    footnote(i = 1, j = 3:6,value = as_paragraph(c("Minor, B>BTar","Moderate, BThr<B<BTar","High, BLim<B<BThr","Major, B<BLim")), ref_symbols = 1:4,part = "header",inline = TRUE)%>%
    fontsize(size=7.5, part='footer')

```

<br>
For the base case scenario and the F reference points, the Probability for each Consequence level used in the Weight of Evidence assessment is shown in Table A.11.4.
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 9. JABBA CPUE_Current.f_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 9. JABBA CPUE_Current.f_Other.sp.csv',sep=''))
Tab=full_join(Tab1,Tab2,by=c("Range","Scenario","finyear","Model"))%>%
  filter(Scenario=='S1')%>%
  dplyr::select(-Scenario)%>%
  rename(Consequence=Range)%>%
  mutate(Consequence=case_when(Consequence=='>lim'~'Major',
                               Consequence=='lim.thr'~'High',
                               Consequence=='thr.tar'~'Moderate',
                               Consequence=='<tar'~'Minor'),
         Consequence=factor(Consequence,levels=c('Minor','Moderate','High','Major')))%>%
  gather(Species,Probability,-c(Consequence,finyear,Model))%>%
  spread(Consequence,Probability)%>%
  mutate(Species=gsub('\\.', ' ', Species))%>%
  rename(Year=finyear)%>%
  mutate(Year=as.character(Year))%>%
  arrange(Model,Species,Year)%>%
  dplyr::select(-Model)
  
 Tab%>% 
    flextable()%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=2,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    set_caption(caption = "Table A.11.4. F reference points. Probability for each consequence levels used in the Weight of Evidence assessment.",
                autonum = NULL)%>%
   footnote(i = 1, j = 3:6,value = as_paragraph(c("Minor, F<FTar","Moderate, FTar<F<FThr","High, FThr<F<FLim","Major, F>FLim.")), ref_symbols = 1:4,part = "header",inline = TRUE)%>%
    fontsize(size=7.5, part='footer')

```

<br>


# Appendix 12. Integrated population dynamics modelling   
We built sex- and age-structured models using Stock Synthesis and time series of catch and available time series of abundance, catch length composition and mean body weight. For sandbar shark, point estimates of F derived from mark-recapture [@McAuley2007b] were also included. The *r4ss* [@taylorVisualizingCatchatageModels2021] and *ss3diags* R packages were used for pre- and post-processing Stock Synthesis inputs and outputs. A full description of Stock Synthesis is available in Methot et al. [-@methotStockSynthesisBiological2013]. Below is a description of model inputs and assumptions.

Life history inputs (M~*a*~, fecundity, maturity, reproductive cycle length, growth, size at birth, length-weight) were sourced from the demographic and SSS modelling described above (Tables A.9.1, A.10.3, Figure A.12.1). Growth at age was assumed to follow the von Bertalanffy growth relationship. Mean length at age was assumed to follow a normal distribution with juvenile and adult CVs set at 0.14 and 0.05 for long-lived species and 0.1 for short-lived species, respectively, based on the age-length data dispersion observed for gummy, dusky, whiskery and sandbar sharks. Fecundity was assumed to increase linearly with maternal length. For species when only a litter size range was reported, the intercept of the linear relationship was prespecified at the litter size mean and the slope was set at 0. The fecundity parameters were divided by the mean of the reproductive cycle length to calculate annual pup production. We assumed a Beverton-Holt stock-recruitment relationship parametrised as a function of lnRo, *h* and the standard deviation in recruitment ($\sigma_R$). Unexploited equilibrium recruitment was estimated whereas *h* was prespecified at the mean value used for SSS modelling (Table A.10.3). Deviates from the stock-recruit relationship were prespecified following examination of preliminary Stock Synthesis outputs with the *r4ss* package (e.g., Courtney et al. [-@courtneyStockSynthesisSS32017]). Preliminary examination with the *r4ss* package also allowed adjusting expected recruitment bias so that resulting recruitment on the standard scale is mean unbiased [@methotAdjustingBiasDue2011]. 

The start and end years of the model were 1941 and 2021, respectively. Initial fishing mortality was not estimated as fishing was assumed to be negligible prior to 1941 [@Braccini2021]. Available time series of catch, abundance, length composition and mean body weight data (Figure A.12.2) were assigned to 'fleets' and 'surveys' (Table A.12.1). For dusky shark, abundance series were not considered for reasons explained in Appendix 11 so a catch-length model was constructed [@Rudd2021]. Catch-length models were also built for smooth hammerhead and spinner shark because abundance time series were not available for smooth hammerhead and available TDGDLF CPUE was not considered to represent abundance trend for spinner shark as explained in Appendix 11.

Gear selectivity estimates for the main fleets capturing the assessed species were only available or could be estimated for dusky, gummy, milk, sandbar, spinner and whiskery sharks, and smooth hammerhead. Hence, Stock Synthesis models were built for these species only. Logistic and dome-shaped selectivity functions were assumed depending on the characteristics of the different fleets and species. For the southern shark fisheries, which use predominately gillnets, dome-shaped selectivity parameters were mostly estimated using the empirical estimates derived by Kirkwood and Walker [-@Kirkwood1986a], Simpfendorfer and Unsworth [-@Simpfendorfer1998c], McAuley et al. [-@McAuley2007a] and Braccini et al. [-@Braccini2022], as initial values although for spinner shark, due to data limitations, they were mostly prespecified to the values reported by Braccini et al. [-@Braccini2022] for the family Carcharhinidae. For the northern shark fisheries, which used predominately hooks, (asymptotic) logistic selectivity parameters were estimated for milk shark or prespecified at 50% length at maturity for spinner shark, as mostly adult individuals were caught by these fisheries. For the other species taken in the northern shark fisheries, dome-shaped selectivity was assumed given their large size and capacity to bend hooks, which was reflected in the small number of large individuals in the length composition. For the Indonesian, Taiwanese and other fleets, selectivity was mirrored to the northern shark fisheries selectivity. For dusky shark, given the current maximum retention size limit (70 cm inter dorsal length), a retention function was implemented with discard mortality pre-specified at 60% [@bracciniHighPostcaptureSurvival2012]. Sex-specific length composition from the northern and southern shark fisheries were collected by onboard observers as fork length (FL) [@Braccini2016b]. This was input into Stock Synthesis as total length (TL) applying published conversion factors [@bartesLengthLengthRelationships2023]. For each fleet, only years with at least 150 observations were included. The TL data were compiled in length bins of 5 cm. Due to the limited number of observations by sex, male and female length composition was combined for milk shark for the northern shark fisheries and the survey, for dusky shark for the survey, and for smooth hammerhead for the southern shark fisheries. Male selectivity was estimated as a female selectivity offset for sandbar shark for the northern shark fisheries and for whiskery and gummy sharks for the southern shark fisheries. For other species/fleets, males and females were assumed to have equal selectivity. Time-varying selectivity, using time blocks (before and after 1999), was implemented for sandbar sharks captured in the southern shark fisheries due to changes in the spatial distribution of length composition samples, particularly in the West Coast (Appendix 13). For whiskery and gummy sharks, a male offset was used in the estimation of selectivity parameters for the southern shark fisheries due to their spatial sexual segregation [@Braccini2016b].

Following Francis [-@Francis2011], we employed the two-stage data weighting approach available in SS3 to iteratively tune variance adjustment factors for fleet-specific abundance indices externally to the model (Stage 1) and length composition data within SS3 (Stage 2). For the abundance indices, the same weighting process applied for building the JABBA models was applied (see Appendix 11). For the length composition data (Stage 2), the *r4ss* package was used to estimate the effective sample size of each data set based on Francis' [-@Francis2011] method and the residuals of the Stock Synthesis model fit (e.g., Courtney et al. [-@courtneyStockSynthesisSS32017]).

Following Carvalho et al. [-@carvalhoCookbookUsingModel2021], model properties were evaluated based on convergence, goodness-of-fit, model consistency and prediction skills as implemented in the *ss3diags* package. Convergence was evaluated based on the objective function final gradient, estimable parameters hitting boundaries, whether the Hessian inverted and by ‘jittering’ the initial value of lnRo. Goodness-of-fit was evaluated based on residual patterns and results of the runs tests. Model consistency was evaluated based on likelihood profiles and retrospective analysis. Finally, prediction skill was evaluated based on hindcasting. Runs tests, hindcasting and retrospective analyses were done only for species with a continuous series of available abundance data in recent years. 

Uncertainty in estimated and derived parameters was obtained from asymptotic standard errors. Finally, sensitivity tests were done to evaluate the effect of the assumed *h* and M~*a*~ values (Table A.12.1).


<!---BLOCK_LANDSCAPE_START--->
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 10. Age.based_SS_scenarios_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 10. Age.based_SS_scenarios_Other.sp.csv',sep=''))
SP_ind=Tab1%>%pull(Species)
SP_other=Tab2%>%pull(Species)
Species=unique(sort(c(SP_ind,SP_other)))

Tab=rbind(Tab1,Tab2)%>%
  arrange(Species)%>%
  mutate(Fem.awt=formatC(Fem.awt, format = "e", digits = 1),
         Mal.awt=formatC(Mal.awt, format = "e", digits = 1),
         Fem.bwt=formatC(Fem.bwt, digits = 2),
         Mal.bwt=formatC(Mal.bwt, digits = 2),
         Fem.fec.a=formatC(Fem.fec.a, format = "e", digits = 1),
         Fem.fec.b=formatC(Fem.fec.b, digits = 2),
         Ln_R0_init=formatC(Ln_R0_init, digits = 1),
         Fem.K=formatC(Fem.K, digits = 2),
         Mal.K=formatC(Mal.K, digits = 2),
         a_FL.to.TL=formatC(a_FL.to.TL, digits = 2),
         b_FL.to.TL=formatC(b_FL.to.TL, digits = 2),
         h.mean=round(h.mean,2),
         SR_sigmaR=round(SR_sigmaR,2))%>%
  dplyr::select(-c(F_ballpark,M.at.age,extra.SD.Q,Msd,h.sd,Mmean,NSF.selectivity))%>%
  relocate(Ln_R0_init,.before=Ln_R0_max)%>%
  relocate(SR_sigmaR,.after=h.mean)

if('Daily.cpues'%in%names(Tab))
{
  Tab=Tab%>%mutate(Daily.cpues=ifelse(is.na(Daily.cpues),'',Daily.cpues))
  flextable(Tab)%>%
  fontsize(size=7, part='body')%>%
  fontsize(j='Fleets',size=6, part='body')%>%
  fontsize(size=8, part='header')%>%
  width(width=.9,unit='cm')%>%
  width(j='Species',width=1.9,unit='cm')%>%
  width(j='Scenario',width=1.4,unit='cm')%>%
  width(j='Fleets',width=2.2,unit='cm')%>%
  width(j='Daily.cpues',width=1.7,unit='cm')%>%
  width(j='Fem.awt',width=1.4,unit='cm')%>%
  width(j='Mal.awt',width=1.4,unit='cm')%>%
  width(j='Fem.TL50',width=1,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=c('Species','Fleets'),align='left',part='all')%>%
  set_header_labels(
      Species="",Scenario="",h.mean="",SR_sigmaR="",Ln_R0_min="Low",Ln_R0_init="Init.",Ln_R0_max="Up",Daily.cpues="CPUE",TLo="",
      Fem.TL.inf="TL.inf",Fem.K="k",Fem.awt="a",Fem.bwt="b",
      Fem.TL50="TL50",Fem.fec.a="f_a",Fem.fec.b="f_b",
      Mal.TL.inf="TL.inf",Mal.K="k",Mal.awt="a",Mal.bwt="b",
      a_FL.to.TL="",b_FL.to.TL="",Fleets="")%>%
  add_header_row(values = c("Species","Scenario","h","SigmaR","lnRo","Removed","TLo","Female","Male","L_a","L_b","Fleets"),
                 colwidths = c(1,1,1,1,3,1,1,7,4,1,1,1))%>%
  flextable::compose(i=1,j=3,part='header',value=as_paragraph(as_i('h')))%>%
  flextable::compose(i=1,j=4,part='header',value=as_paragraph(stri_unescape_unicode('\u03c3'),as_sub('R')))%>%
  flextable::compose(i=2,j='Fem.TL50',part='header',value=as_paragraph('TL',as_sub('50')))%>%
  flextable::compose(i=2,j=c(10,17),part='header',value=as_paragraph('TL',as_sub(stri_unescape_unicode('\\u221E'))))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.12.1. Input parameter values for the integrated population dynamics models. h, steepness; σR, recruitment standard deviation; lnRo, log of unexploited equilibrium recruitment; Low, lower bound; Init., initial value; Up, upper bound; TLo, total length at birth (cm); TL50, total length at 50% maturity (cm); k, growth coefficient; a and b, parameters of the length-weight relationship (in cm and kg, respectively); f_a and f_b, parameters of the fecundity at age relationship; L_a and L_b, parameters of the TL-FL (fork length, in cm) relationship.",
              autonum = NULL)
}else
{
  flextable(Tab)%>%
  fontsize(size=7, part='body')%>%
  fontsize(j='Fleets',size=6, part='body')%>%
  fontsize(size=8, part='header')%>%
  width(width=.9,unit='cm')%>%
  width(j='Species',width=1.8,unit='cm')%>%
  width(j='Scenario',width=1.5,unit='cm')%>%
  width(j='Fleets',width=2.2,unit='cm')%>%
  width(j='Fem.awt',width=1.4,unit='cm')%>%
  width(j='Mal.awt',width=1.4,unit='cm')%>%
    width(j='Fem.fec.a',width=1.3,unit='cm')%>%
  width(j='Fem.TL50',width=1,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=c('Species','Fleets'),align='left',part='all')%>%
  set_header_labels(
      Species="",Scenario="",h.mean="",SR_sigmaR="",Ln_R0_min="Low",Ln_R0_init="Init.",Ln_R0_max="Up",TLo="",
      Fem.TL.inf="TL.inf",Fem.K="k",Fem.awt="a",Fem.bwt="b",
      Fem.TL50="TL50",Fem.fec.a="f_a",Fem.fec.b="f_b",
      Mal.TL.inf="TL.inf",Mal.K="k",Mal.awt="a",Mal.bwt="b",
      a_FL.to.TL="",b_FL.to.TL="",Fleets="")%>%
  add_header_row(values = c("Species","Scenario","h","SigmaR","lnRo","TLo","Female","Male","L_a","L_b","Fleets"),
                 colwidths = c(1,1,1,1,3,1,7,4,1,1,1))%>%
  flextable::compose(i=1,j=3,part='header',value=as_paragraph(as_i('h')))%>%
  flextable::compose(i=1,j=4,part='header',value=as_paragraph(stri_unescape_unicode('\u03c3'),as_sub('R')))%>%
  flextable::compose(i=2,j='Fem.TL50',part='header',value=as_paragraph('TL',as_sub('50')))%>%
  flextable::compose(i=2,j=c(9,16),part='header',value=as_paragraph('TL',as_sub(stri_unescape_unicode('\\u221E'))))%>%
  border_inner_h(border = fp_border_default(width = 0), part = "header")%>%
  set_caption(caption = "Table A.12.1. Input parameter values for the integrated population dynamics models. h, steepness; σR, recruitment standard deviation; lnRo, log of unexploited equilibrium recruitment; Low, lower bound; Init., initial value; Up, upper bound; TLo, total length at birth (cm); TL50, total length at 50% maturity (cm); k, growth coefficient; a and b, parameters of the length-weight relationship (in cm and kg, respectively); f_a and f_b, parameters of the fecundity at age relationship; L_a and L_b, parameters of the TL-FL (fork length, in cm) relationship.",
              autonum = NULL)%>%
   footnote(i = 1, j = 22,value = as_paragraph("Northern.shark, Northern Shark Fisheries; Southern.shark_1, TDGDLF monthly records; Southern.shark_2, TDGDLF daily records; WRL, West Coast Rock lobster Fishery"), ref_symbols = 'a',part = "header",inline = TRUE)%>%
   fontsize(size=7.5, part='footer')
}
  

```
<!---BLOCK_LANDSCAPE_STOP--->

<!-- Display biological relationships -->
```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.1.',i,'. ',Titl,'. Biological relationships. FL, fork length; TL, total length; TLMax, Maximum total length; TL95 and TL50, total length at 95% and 50% maturity; TL0, TL at birth; Age50; age at 50% maturity (years; range provided); AgeMax, maximum age (years; range provided).',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           '1_Inputs/Visualise data/Biology.tiff',sep='/'),sep='')), sep = '\n')
  cat("\n")
}
```
<!-- Display Selectivity at length for S1--> 
The modelled selectivity at lengths for each fleet is shown in Figure A.12.2.

```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
  Titl=gsub(".*\\.", "", Species[i])
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                                'SS3 integrated/S1/plots',sep='/'), sep=''))
  x.indx=FilEs[grep('sel01_multiple_fleets_length1',FilEs)]
  if(length(x.indx)>0)
  {
    for(x in 1:length(x.indx))
    {
      cat(sprintf('![%s](%s)', paste('Figure A.12.2.',i,'. ',Titl,'. Selectivity at length by fleet.',sep=''), 
                  paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                               'SS3 integrated/S1/plots/',sep='/'),x.indx[x], sep='')), sep = '\n')
      cat("\n")
    }
    ii=ii+1
  }
}
```
<br> 

<br>

All available data time series for each of the assessed species is shown in Figure A.12.3. For the current assessment, only catch series, size composition series for the TDGDLF, NSF and Survey, mean weight series for the TDGDLF, Fishing mortality series for the TDGDLF and abundance series were used. Other available time series, such as conventional and acoustic tagging, are also displayed for consideration in future assessments. 


<!-- Display available time series --> 
```{r, results='asis', fig.show='asis',fig.align='left',fig.height=6,fig.width=7}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.3.',i,'. ',Titl,'. Available data time series.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           '1_Inputs/Visualise data/avail.dat.tiff',sep='/'),sep='')), sep = '\n')
  cat("\n")
}
```

<br>
<!-- Display Sensitivity analysis--> 
The resulting relative (spawning) biomass trajectories based on the scenarios considered are shown in Figures A.12.4 for indicator species and A.12.5 for non-indicator species. Overall, the different scenarios considered showed very similar trajectories and current depletion levels. On the basis of these comparisons, S1 was selected as the base case scenario.

<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_SS3 integrated_Indicator.sp_sensitivity.tiff',sep='')
```
![Figure A.12.4. Sensitivity analysis showing the influence of the assumed h (S2) and Ma (S3) for indicator species on the estimated relative (spawning) biomass (i.e. biomass over unfished biomass, solid line) with 95% confidence intervals (shaded area). Also shown are the reference points (red, limit; orange, threshold; green, target). The assumed h and Ma values for the different scenarios is shown in Table A.12.1. ](`r path`){width=800px}


```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_SS3 integrated_Other.sp_sensitivity.tiff',sep='')
```
![Figure A.12.5. Sensitivity analysis showing the influence of the assumed h (S2) and Ma (S3) for non-indicator species on the estimated relative (spawning) biomass (i.e. biomass over unfished biomass, solid line) with 95% confidence intervals (shaded area). Also shown are the reference points (red, limit; orange, threshold; green, target). The assumed h and Ma values for the different scenarios is shown in Table A.12.1.](`r path`){width=800px}

<br>

<!-- Display Fit to cpue  and length series for S1--> 
For the base case scenario of all assessed species, the model final gradient was small (<$1e^{-4}$), the Hessian was positive definite and parameters were not estimated close to bounds (Table A.12.2). The models typically provided reasonably good visual fits to abundance series, with most of the estimated mean CPUE values lying within the 95% CI (Figure A.12.6). The models provided moderate or good visual fits to the available length composition data and mean body weight data (Figures A.12.7 and  A.12.8). The runs tests provided no evidence to reject the hypothesis of randomly distributed residuals associated with fits to abundance series (Figure A.12.9). Runs tests were not done for dusky and spinner sharks, and smooth hammerhead due to data limitations (only sporadic length compositions were available, mainly from earlier years). The goodness-of-fit was assessed as adequate for the abundance data (RMSE≤30%, [@carvalhoCookbookUsingModel2021]) but for most species, length composition showed larger RMSE values and residual boxplots, and an apparent trend (Figure A.12.10). 

<br>

```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
  Titl=gsub(".*\\.", "", Species[i])
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/plots',sep='/'), sep=''))
  x.indx=FilEs[grep('index2',FilEs)]
  if(length(x.indx)>0)
  {
    for(x in 1:length(x.indx))
    {
      xxx=x.indx[x]
      if(grepl('F.series',xxx)) xxx=str_replace(str_replace(xxx, "F.series","F series"),".png","")
        cat(sprintf('![%s](%s)', paste('Figure A.12.6.',i,'. ',paste(Titl,str_replace(str_replace(xxx, "index2_cpuefit_",""),".png",""),sep='. '),'. Fits to the abundance series for the base case scenario. Observed (white circles and 95% CIs) and predicted (solid blue line) abundance.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/plots/',sep='/'),x.indx[x], sep='')), sep = '\n')
  cat("\n")
    }
    ii=ii+1
  }
}
```

<br>

```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
  Titl=gsub(".*\\.", "", Species[i])
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/plots',sep='/'), sep=''))
  x.indx=FilEs[grep('comp_lenfit__aggregated_across',FilEs)]
  if(length(x.indx)>0)
  {
    for(x in 1:length(x.indx))
    {
        cat(sprintf('![%s](%s)', paste('Figure A.12.7.',i,'. ',Titl,'. Fits to the length composition series for the base case scenario. Observed (grey-shaded area) and predicted (solid line) composition.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/plots/',sep='/'),x.indx[x], sep='')), sep = '\n')
  cat("\n")
    }
    ii=ii+1
  }
}
```

<br>

```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
  Titl=gsub(".*\\.", "", Species[i])
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/plots',sep='/'), sep=''))
  x.indx=FilEs[grep('bodywt_fit_flt',FilEs)]
  if(length(x.indx)>0)
  {
    for(x in 1:length(x.indx))
    {
        cat(sprintf('![%s](%s)', paste('Figure A.12.8.',ii,'. ',Titl,'. Fits to the mean body weight series for the base case scenario. Observed (white circles and 95% CIs) and predicted (solid blue line) mean body weight.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/plots/',sep='/'),x.indx[x], sep='')), sep = '\n')
  cat("\n")
    }
    ii=ii+1
  }
}
```

<br>

<!-- Display Runs test for S1--> 
```{r, results='asis'}
ii=1
for(i in 1:length(Species)) 
{
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/Diagnostics',sep='/'), sep=''))
  x.indx=FilEs[grep('runs_tests_cpu',FilEs)]
  if(length(x.indx)>0)
  {
      Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.9.',ii,'. ',Titl,'. Runs test for the base case scenario and fishery. Green panel, no evidence of lack of randomness of abundance series residuals (p≥0.05); Red panel, evidence of lack of randomness of abundance series residuals. The shaded area spans three standard errors and red circles denote years with greater residuals.',sep=''),
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,
                                           'SS3 integrated/S1/Diagnostics/runs_tests_cpue.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
  ii=ii+1
  }
}
```

<br>

<!-- Display dresiduals for S1--> 
```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.10.',i,'. ',Titl,'. Residual diagnostic plots for the base case scenario. Boxplot, median and quantiles of the residuals; solid black line, loess smoother through the residuals.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/Diagnostics/jabbaresidual.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!-- Display Likelihood profiles for S1--> 
Likelihood profiles allow the identification of the relative information (and tensions among data types) contained in each data type and they provide a sense of the likelihood surface surrounding maximum likelihood estimates (MLEs). For this, a parameter of interest is fixed at different values and the model is re-estimated. The equilibrium recruitment parameter, Ro, is commonly profiled because it represents an ideal global scaling parameter given that unfished (virgin) recruitment is proportional to unfished biomass [@carvalhoCookbookUsingModel2021]. 

For each species and the base case scenario, the Ro likelihood profiles are shown in the Figure A.12.11 series. With the exception of spinner shark, there was no clear global minima for the total likelihood. For all species, the total negative log-likelihood showed a large decrease along the profile from low Ro values to the MLE values. For Ro values larger than the MLE, the likelihood profile showed a more gradual increase for most species except for gummy and dusky sharks, which showed no increase or a slight increase, respectively. This suggests that the available data sources for the models for these species are relatively uninformative, highlighting the data-limited nature of these assessments. For dusky, gummy, milk, sandbar and whiskery sharks, the likelihood profile for the penalty on the recruitment deviations had the greatest likelihood profile gradient. For smooth hammerhead and spinner sharks, length composition had the greatest likelihood profile gradient.   

<br>
  
```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
   FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/Diagnostics',sep='/'), sep=''))
  x.indx=FilEs[grep('R0_profile_plot',FilEs)]
  if(length(x.indx)>0)
  {
     Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.11.',ii,'. ',Titl,'. Log-likelihood profiles for Ro for the various data components included in the models. The vertical broken lines indicate the estimated value.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/Diagnostics/R0_profile_plot.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
    ii=ii+1
  }
 
}
```


<!-- Display Retrospective analysis for S1-->  
 For species with abundance series and the base case scenario, there was an overall negligible retrospective pattern, producing consistent stock status estimates for milk, sandbar, and whiskery sharks (Figure A.12.12) and therefore supporting the use of the base case model for making future projections. For sandbar shark, the Mohn's-rho (ρ) statistic fell within the acceptable interval for long-lived species of -0.15 and 0.2 whereas for whiskery and milk sharks, the ρ statistic fell within the acceptable interval for short-lived species of -0.22 and 0.3 [@carvalhoCookbookUsingModel2021]. For gummy shark, however, ρ (0.37) fell outside the expected ranges of either short- or long-lived species and the 2016 peel fell outside the upper 95% CI for stock fecundity, yielding a positive retrospective pattern, i.e. an overestimation of biomass. This indicates that data at least up to 2017 are needed to produce stable results. Retrospective analyses were not done for dusky and spinner sharks, and smooth hammerhead due to data limitations (only sporadic length compositions were available, mainly from earlier years).
<br>
  
```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
   FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/Diagnostics',sep='/'), sep=''))
  x.indx=FilEs[grep('retro_Mohns_Rho',FilEs)]
  if(length(x.indx)>0)
  {
     Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.12.',ii,'. ',Titl,'. Retrospective analysis results for the base case scenario done by removing one year at a time sequentially (n=5) and predicting relevant management quantities. Grey shaded area, 95% CI from the reference model.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/Diagnostics/retro_Mohns_Rho.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
    ii=ii+1
  }
 
}
```

<!-- Display Hindcasting for S1--> 
For the base case scenario, the hindcasting cross-validation results showed predictions fell within the 95% CI suggesting good predicting skills overall (Figure A.12.13). For gummy shark, the mean absolute scale error (MASE) score was > 1 (the reference level), which implies that the average model forecast is no better than a random walk [@carvalhoCookbookUsingModel2021]. For the other species, the MASE score was ≤ 1 (i.e., predictions are better than a random walk). Hindcasting was not done for dusky and spinner sharks, and smooth hammerhead due to data limitations (only length composition was available from sporadic earlier years mostly).
<br>
  
```{r, results='asis'}
ii=1
for(i in 1:length(Species))
{
  FilEs=list.files(paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                             'SS3 integrated/S1/Diagnostics',sep='/'), sep=''))
  x.indx=FilEs[grep('Hindcasting cross-validation_Survey',FilEs)]
  if(length(x.indx)>0)
  {
    Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.13.',i,'. ',Titl,'. Hindcasting cross-validation results for the base case scenario showing one-year ahead forecast of abundance values, performed with 5 hindcast model runs relative to the expected value. Large color-coded circles, observations; grey shaded area, predicted 95% CI for the reference year; small color-coded circles, model predictions.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/Diagnostics/Hindcasting cross-validation_Survey.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
  ii=ii+1
  }
  
}
```

<br>

<!-- Display Jitters for S1--> 
For the base case scenario, the jitter test for global convergence is shown in Figure A.12.14. For smooth hammerhead and spinner shark, all of the 50 iterations of the jitter test converged at the total likelihood estimate value of the base model run. For the other species, most of the iterations converged at or close to the total likelihood estimate value of the base model run with 3, 2, 8, 6, 2 converging at greater values for dusky, gummy, milk, sandbar and whiskery sharks, respectively. This suggests that the jittered models were some how sensitive to the initial values of the parameters. However, all converged model runs resulted in total likelihood values equal to or greater than the base model hence the jitter test supports the hypothesis that the base model optimisation converged to a global solution. 
<br>
  
```{r, results='asis'}
for(i in 1:length(Species)){
  Titl=gsub(".*\\.", "", Species[i])
  cat(sprintf('![%s](%s)', paste('Figure A.12.14.',i,'. ',Titl,'. Jitter plots for the base case scenario. Top panel, Spawning stock biomass trajectories obtained from the jittered model runs. Gray shaded areas are the 95% confidence intervals from the base case run. Bottom panel, solid black circles represent the total likelihood obtained from jittered model runs. The red horizontal line represents the total likelihood obtained from the base case run.',sep=''), 
              paste(Model.outputs.an,paste(paste('1.',Species[i],sep=''),this.yr,  
                                           'SS3 integrated/S1/Diagnostics/Jitter.tiff',sep='/'), sep='')), sep = '\n')
  cat("\n")
}
```

<br>

<!---BLOCK_LANDSCAPE_START--->
<!-- Display parameter estimates for S1--> 
  For the base case scenario, model gradient and parameter estimates are shown in Table A.12.2 and management quantities of interest are shown in Table A.12.3.
```{r,echo=FALSE,ft.align="left"}
#parameters
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 11. Age.based_SS_estimates_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 11. Age.based_SS_estimates_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
  filter(Scenario=='S1')%>%
  filter(!Par%in%c('F.target','F.threshold','F.limit'))%>%
  arrange(Species)%>%
  dplyr::select(Species,Par,Value,Phase,Min,Max,Init,Status,Parm_StDev,Gradient)%>%
  mutate_at(vars(c('Value','Min','Max','Init','Parm_StDev')), round, 2)

dis.sp=unique(Tab$Species)
Model.gradient=data.frame(Species=dis.sp,
                          Model.Gradient=NA)
for(i in 1:length(dis.sp))
{
  this.wd=paste(handl_OneDrive("Analyses/Population dynamics/1."),dis.sp[i],"/",this.yr,"/SS3 integrated",sep='')
  this.wd1=paste(this.wd,"S1",sep='/')
  MLE=read.admbFit(paste(this.wd1,'ss',sep='/'))
  Model.gradient$Model.Gradient[i]=MLE$maxgrad
}

Tab=Tab%>%
  left_join(Model.gradient,by='Species')%>%
  mutate(Gradient=format(Gradient, scientific=T,digits=2),
         Model.Gradient=format(Model.Gradient, scientific=T,digits=2))%>%
  relocate(Model.Gradient,.after=Species)%>%
  relocate(Init,.after=Min)%>%
  relocate(Value,.before = Parm_StDev)%>%
  rename('Model Gradient'=Model.Gradient,
         Estimate=Value,
         SE=Parm_StDev)%>%
  dplyr::select(-Gradient)
duplis=duplicated(paste(Tab$Species,Tab$'Model Gradient'))
Tab$'Model Gradient'[duplis] <- ""
Tab%>%
  flextable()%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=2.5,unit='cm')%>%
  width(j=c('Phase','Min','Max','Init','Status'),width=1.5,unit='cm')%>%
  width(j=c('Par'),width=6,unit='cm')%>%
  width(j=c('SE','Model Gradient'),width=2.5,unit='cm')%>% 
  align(align='center',part='all')%>%
  align(j=c('Species','Par'),align='left',part='all')%>%
  set_caption(caption = "Table A.12.2. Summary of parameter estimates.",
              autonum = NULL)
```
<!---BLOCK_LANDSCAPE_STOP--->
<br>

```{r,echo=FALSE,ft.align="left"}
#quantities of interest
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 11. Age.based_SS_quantities_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 11. Age.based_SS_quantities_Other.sp.csv',sep=''))
Tab=rbind(Tab1,Tab2)%>%
  filter(Scenario=='S1')%>%
  arrange(Species)%>%
  mutate(SE=formatC(SE, format = "e", digits = 2))%>%
  mutate_at(vars('Median'), round, 2)%>%
  dplyr::select(-Scenario)

Tab%>%
  flextable()%>%
  fontsize(size=8, part='body')%>%
  fontsize(size=9, part='header')%>%
  width(width=2.5,unit='cm')%>%
  align(align='center',part='all')%>%
  align(j=c('Species','Label'),align='left',part='all')%>%
  width(j='Label',width=3.5,unit='cm')%>%
  set_header_labels(Label = "Quantity")%>%
  set_caption(caption = "Table A.12.3. Management quantities of interest for the base case scenario.",
              autonum = NULL)


```
<br> 

<!-- Display biomass trajectories for S1--> 
For the base case scenario, the relative (spawning) biomass trajectories for indicator species are shown in Figure A.12.15.
<br>

```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_SS3 integrated_Indicator.sp.tiff',sep='')
```
![Figure A.12.15. Estimated time series of relative (spawning) biomass (i.e. biomass over unfished biomass) for indicator species (60% and 95% confidence intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target) and the total catch series (broken blue line).](`r path`){width=800px}

For the base case scenario, the relative (spawning) biomass trajectories for non-indicator species are shown in Figure A.12.16.
<br>
  
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Relative.biomass_SS3 integrated_Other.sp.tiff',sep='')
```
![Figure A.12.16. Estimated time series of relative (spawning) biomass (i.e. biomass over unfished biomass) for non-indicator species (60% and 95% confidence intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target) and the total catch series (broken blue line).](`r path`){width=800px}

<br>

<!-- Display F trajectories for S1--> 
For the base case scenario, the F trajectories for indicator species are shown in Figure A.12.17.
<br>
  
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.series_SS3 integrated_Indicator.sp.tiff',sep='')
```
![Figure A.12.17. Estimated time series of fishing mortality for indicator species (60% and 95% confidence intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

For the base case scenario, the F trajectories for non-indicator species are shown in Figure A.12.18.
<br>
  
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/F.series_SS3 integrated_Other.sp.tiff',sep='')
```
![Figure A.12.18. Estimated time series of fishing mortality for non-indicator species (60% and 95% confidence intervals shown in dark grey and light grey, respectively). Forward projections are highlighted in brown. Also shown are the reference points (red, limit; orange, threshold; green, target).](`r path`){width=800px}

<br>

<!-- Display Kobe plots for S1--> 
For the base case scenario, Kobe plots for indicator species are shown in Figure A.12.19.
<br>
  
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_SS3 integrated_WA_Fisheries_Indicator.sp.tiff',sep='')
```
![Figure A.12.19. Kobe plots showing the estimated trajectories of relative (spawning) biomass (i.e. biomass over unfished biomass, B0) and fishing mortality for indicator species. The red, orange, yellow and green panels indicate the areas below (above for fishing mortality) limit, between limit and threshold, between threshold and target, and above (below for fishing mortality) reference points.](`r path`){width=800px}

For the base case scenario, Kobe plots for non-indicator species are shown in Figure A.12.20.
<br>
  
```{r, echo=FALSE}
path=paste(Model.outputs.RAR,'/Kobe_plot_SS3 integrated_WA_Fisheries_Other.sp.tiff',sep='')
```
![Figure A.12.20. Kobe plots showing the estimated trajectories of relative (spawning) biomass (i.e. biomass over unfished biomass, B0) and fishing mortality for non-indicator species. The red, orange, yellow and green panels indicate the areas below (above for fishing mortality) limit, between limit and threshold, between threshold and target, and above (below for fishing mortality) reference points.](`r path`){width=800px}

<br>

<!-- Display likelihoods for each Consequence--> 
For the base case scenario and the biomass reference points, the Probability for each Consequence level used in the Weight of Evidence assessment is shown in Table A.12.4.
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 12. Age.based_SS_current.depletion_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 12. Age.based_SS_current.depletion_Other.sp.csv',sep=''))
Tab=full_join(Tab1,Tab2,by=c("Range","Scenario","finyear","Model"))%>%
  filter(Scenario=='S1')%>%
  dplyr::select(-Scenario)%>%
  rename(Consequence=Range)%>%
  mutate(Consequence=case_when(Consequence=='<lim'~'Major',
                               Consequence=='lim.thr'~'High',
                               Consequence=='thr.tar'~'Moderate',
                               Consequence=='>tar'~'Minor'),
         Consequence=factor(Consequence,levels=c('Minor','Moderate','High','Major')))%>%
  gather(Species,Probability,-c(Consequence,finyear,Model))%>%
  spread(Consequence,Probability)%>%
  mutate(Species=gsub('\\.', ' ', Species))%>%
  rename(Year=finyear)%>%
  mutate(Year=as.character(Year))%>%
  arrange(Species,Year)%>%
  dplyr::select(-Model)
  
 Tab%>% 
    flextable()%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=2,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    set_caption(caption = "Table A.12.4. Biomass reference points. Probability for each consequence levels used in the Weight of Evidence assessment.",
                autonum = NULL)%>%
   footnote(i = 1, j = 3:6,value = as_paragraph(c("Minor, B>BTar","Moderate, BThr<B<BTar","High, BLim<B<BThr","Major, B<BLim")), ref_symbols = 1:4,part = "header",inline = TRUE)%>%
   fontsize(size=7.5, part='footer')

```

<br>  

For the base case scenario and the F reference points, the Probability for each Consequence level used in the Weight of Evidence assessment is shown in Table A.12.5.
```{r,echo=FALSE,ft.align="left"}
Tab1<- read.csv(paste(Model.outputs.RAR,'/Table 12. Age.based_SS_Current.f_Indicator.sp.csv',sep=''))
Tab2<- read.csv(paste(Model.outputs.RAR,'/Table 12. Age.based_SS_Current.f_Other.sp.csv',sep=''))
Tab=full_join(Tab1,Tab2,by=c("Range","Scenario","finyear","Model"))%>%
  filter(Scenario=='S1')%>%
  dplyr::select(-Scenario)%>%
  rename(Consequence=Range)%>%
  mutate(Consequence=case_when(Consequence=='>lim'~'Major',
                               Consequence=='lim.thr'~'High',
                               Consequence=='thr.tar'~'Moderate',
                               Consequence=='<tar'~'Minor'),
         Consequence=factor(Consequence,levels=c('Minor','Moderate','High','Major')))%>%
  gather(Species,Probability,-c(Consequence,finyear,Model))%>%
  spread(Consequence,Probability)%>%
  mutate(Species=gsub('\\.', ' ', Species))%>%
  rename(Year=finyear)%>%
  mutate(Year=as.character(Year))%>%
  arrange(Species,Year)%>%
  dplyr::select(-Model)
  
 Tab%>% 
    flextable()%>%
    fontsize(size=8, part='body')%>%
    fontsize(size=9, part='header')%>%
    width(width=2,unit='cm')%>%
    width(j='Species',width=2.5,unit='cm')%>%
    set_caption(caption = "Table A.12.5. F reference points. Probability for each consequence levels used in the Weight of Evidence assessment.",
                autonum = NULL)%>%
   footnote(i = 1, j = 3:6,value = as_paragraph(c("Minor, F<FTar","Moderate, FTar<F<FThr","High, FThr<F<FLim","Major, F>FLim.")), ref_symbols = 1:4,part = "header",inline = TRUE)%>%
   fontsize(size=7.5, part='footer')

```

<br>  

# Appendix 13. Spatio-temporal distribution of available length composition   

<br>
  
```{r, echo=FALSE}
path=handl_OneDrive(paste('Analyses/Size and sex patterns/spatial_length_comp/Pop din_spatial_sandbar shark_West.tiff'))
```
![Figure A.13.1. Spatio-temporal distribution of sandbar shark length composition samples collected in the West coast by boat and mesh size.](`r path`){width=800px}
